---
title: "Bookends 0.4"
output: html_notebook

#  bookdown::word_document2:
#    reference_docx: word_styles_reference_0-2.docx   
---
```{r options, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE)
```


# Conclusion

**Overview**
  
<!-- The diet of a species is a basic aspect of its biology and defines some of its most important ecological relationships. Understanding a species's diet is crucial for successful management (Stier et al. 2016), and may even determine the limitations of recovery (Ferrer & Negro 2004). The diet of individuals can scale up to the population level, with consequences for population growth rates. In this thesis I examined dietary variation and its fitness consequences for northern goshawks in coastal British Columbia. I quantified diet at large (study area) and small (individual nest) scales for breeding goshawks and found that although goshawks consumed a wide variety of prey, consistent with their status as generalist predators, they showed a specialist-like reliance on tree squirrels of the genus *Tamiasciurus*, the extent of which varied from nest to nest. I then examined variation in diet across two ecological zones within my study area, the wetter *coastal zone* and drier *transitional zone* and found that although there were slight differences between the zones, the reliance on tree squirrels remained constant. Finally, I investigated whether two aspects of diet, diversity and specialization on tree squirrels, affected goshawk fitness, but found little evidence to support an affect of diet on productivity. -->

Specialist and generalist predators differ in their degree of dependence on prey species with cascading consequences for many aspects of their life history (Korpimaki & Norrdhal 1991, Resano-Mayor et al. 2016). Specialists are efficient hunters of their main prey at the cost of poor success when hunting other species, whereas generalists hunt many species with equal skill (Terraube et al. 2011). A specialist may struggle to compensate with other prey when its main prey becomes scarce, but generalists readily switch to alternate prey (Steenhof & Kochert 1988, Terrabue & Arroyo 2011). Specialists may even be limited in geographic range or habitat use by the presence and abundance of their main prey (Ferrer & Negro 2004, Peers et al. 2003). As a result, specialists depend on a single species and their demographic parameters--such as migration, reproductive success, and survival--vary in synchrony with its abundance (Korpimaki & Norrdhal 1991, Terraube et al. 2011). In contrast, generalists make use of many species but depend on no single one, and their populations are relatively more stable (Hanski et al. 1991, Andersson & Erlinge 1977). <!-- and are more widespread -->

<!--Korpimaki & Norrdhall 1991: "The ability of nomadic specialists to track population fluctuations of microtines without time lags seems to result from a numerical response based mainly on mobility. In contrast, the numerical responses of resident generalists are based mostly on changes in natality and mortality, causing their response to lag behind the population fluctuations of microtines." -->

<!-- Terraube et al. 2011: "...our study also shows that the specialist had lower foraging success than the closely related generalist species when its main prey was scarce..."

"This helps understand 1) why fewer pallid harriers breed when voles are scarce, and why most fail to breed successfully (Table 1), and 2) the nomadic behaviour of the vole specialist pallid harrier. It might be more efficient to move to a different area rather than to stay in an area where the preferred prey is scarce, even when alternative prey are abundant. -->

This familiar dichotomy between specialist and generalist species is, of course, an oversimplification. The abundance of a single prey species can be a major driver of demographic parameters, such as reproductive success, for a generalist predator as well as for a specialist (Elmhagen et al. 2000, Resano-Mayor et al. 2006). Within a single species some populations (Salamolard et al. 2000, Roth et al. 2007), or some individuals within a population (Woo et al. 2008), may be more or less specialized. Even a single individual may become more specialized over its lifetime as a result of age and experience (Rutz 2006). Correctly identifying the degree of specialization and understanding its effect on demographic parameters is more than a matter of theory or curiosity: the consequences of specialization can scale up from individuals through populations to entire species, with profound implications for conservation (Ferrer & Negro 2004, Terraube et al. 2011, Resano-Mayor 2016).

The complex relationship between dietary specialization and conservation is exemplified by the northern spotted owl. Spotted owls depend on old-growth forests, but the cause of this association has been a source of speculation from the earliest years of spotted owl research (). The association appears to be driven, in part, by the spotted owl's relatively specialized diet (Carey et al. 1992, Ward et al. 1998). More than half the biomass spotted owls consume comes from just two taxa, flying squirrels (*Glaucomys sabrinus*) and woodrats (bushy-tailed woodrat *Neotoma cinerea* and dusky-footed woodrat *N. fuscipes*; reviewed in Carey et al. 1992<!-- but find a newer citation-->). This holds true across the subspecies' range, although the relative contribution of each taxa varies with geographic region and forest type in response to local abundance. In Washington's Olympic Peninsula, where woodrats are absent, spotted owls consume primarily flying squirrels (Carey et al. 1992), whereas in northern California flying squirrels make up a smaller portion of the diet and woodrats, which are more abundant, dominate (Ward et al. 1998). Even within a single spotted owl population some individuals specialize on one taxa or the other (Zabel et al. 1995). Home range sizes in the flying squirrel-dependent Olympic Peninsula are among the largest ever recorded (Carey at al. 1992), and where both taxa are present owls which consume primarily flying squirrels have larger home ranges than those which consume mostly woodrats (Zabel et al. 1995). Evidently diet and prey abundance affect some demographic parameters, such as breeding density, which has led some to recommend increasing prey abundance as a route to increase owl abundance (Forsman et al. 2014). Yet prey abundance does not appear to affect spotted owl productivity (Rosenberg et al. 2003). Instead, productivity appears to be the result of complex interactions between climate and prey abundance (Glenn et al. 2011).<!-- This needs a concluding sentence --><!-- Also needs something about NSPOW pop dynamics -->

<!-- Carey et al 1992; "Six nonexclusive hypotheses have been proposed to explain the requirement of spotted owls for old growth, but only abundance of prey can explain the requirement of Spotted Owls for large areas of old growth." -->

<!-- Ward et al. 1998: "First, we showed that Northern Spotted Owls occurring in mixed evergreen forests of northwestern California selected the dusky-footed woodrat over other prey species. Second, we found that these owls selected certain foraging sites where dusky-footed woodrats were more abundant. [...] the owls. Collectively, these findings provided only a partial explanation for the owl's affinity for late seral forest." -->

<!-- Ward et al. 1998: "Site selection by owls was most pronounced at the ecotone between late and early seral stages where dusky-footed woodrats were most abundant." -->

In contrast to the spotted owl's dependence on a few prey species, the northern goshawk is considered a generalist predator and consumes an enormous diversity of prey across its wide geographic range (reviewed in Drennen 2006). I identified **25** different prey species in the diet of goshawks in coastal British Columbia, which is consistent with a generalist foraging strategy. Some goshawk populations appear to be strongly generalist (Salafsky et al. 2007), whereas in others a key prey species is a major driver of productivity, survival, and other demographic parameters (Doyle & Smith 1994, Tornberg et al. 2005). Over 60% of goshawk diet in my study area was composed of tree squirrels (*Tamiasciurus* spp.), which indicates a level of specialization even greater than that of the spotted owl. However, I found no affect of the degree of dietary specialization on goshawk productivity. There are several explanations for this unexpected finding. First, specialists may not be more productive than generalists: some individual goshawks may prefer tree squirrels and consume them disproportionate to their abundance, but both specialists and generalists are equally fit (Woo et al. 2008). Alternately, specialization may be the result of opportunity: all goshawks take tree squirrels proportionate to their abundance, but it is total prey abundance rather than tree squirrel abundance which drives productivity. Finally, as in the spotted owl, prey abundance and diet during the breeding season may not be a lesser driver of productivity than other factors, such as weather or winter prey abundance.

<!--
Although individual goshawks may specialize to a greater or lesser degree on tree squirrels, specialists may not gain any fitness advantage over generalists (Woo et al. 2008). Alternatively, diet outside the breeding season, or a variable unrelated to diet such as weather, may be a larger driver of productivity than breeding season diet. A final possibility is that the degree of dietary specialization is due not to variable preference for tree squirrels, but rather variable abundance of tree squirrels between territories. Under this scenario, goshawks are generalists which capture tree squirrels proportional to their abundance and the total abundance of all prey, rather than tree squirrel abundance, drives goshawk productivity. Without data on prey abundance, it is unfortunately impossible to test these hypotheses.
-->

```{r echo=FALSE, warning=FALSE, message=FALSE}
# Import conflict settings.
source('../src/conflicted.R')

# Load some libraries.
library(tidyverse)
library(lubridate)
library(sf)
library(adehabitatHR)
library(recurse)
library(flextable)
library(rmapshaper)
library(ggspatial)
library(ggrepel)

# Copy out results from Bloxton.
blx <- tribble(
  ~class, ~family, ~category, ~common, ~prop.count,
  'Aves', 'Phasianidae', 'large bird', 'grouse', 31,
  'Aves', 'Columbidae', 'large bird', 'band-tailed pigeon', 22,
  'Aves', 'Corvidae', 'medium bird', "steller\'s jay", 14,
  'Aves', 'Picidae', 'medium bird', 'woodpecker', 4,
  'Aves', 'Turdidae', 'small bird', 'thrush', 4,
  'Aves', 'Unknown', 'small bird', 'unknown', 2,
  'Mammalia', 'Leporidae', 'large mammal', 'snowshoe hare', 10,
  'Mammalia', 'Sciuridae', 'medium mammal', 'douglas squirrel', 6,
  'Mammalia', 'Sciuridae', 'small mammal', 'flying squirrel', 4,
  'Mammalia', 'Rodentia', 'small mammal', 'unknown', 3
)

# This assumes "other birds" and "other mammals" are "small" sized, and splits woodpeckers and thrushes (which he gives as a combine 8%) into 4% each.

# Copy out results from Thraikill.
thr <- tribble(
  ~class, ~family, ~category, ~common, ~prop.count,
  'Aves', 'Phasianidae', 'large bird', 'ruffed grouse', 45,
  'Aves', 'Corvidae', 'medium bird', "steller\'s jay", 13,
  'Aves', 'Turdidae', 'medium bird', 'american robin', 13,
  'Aves', 'Phasianidae', 'large bird', 'ring-necked pheasant', 8,
  'Aves', 'Odontophoridae', 'large bird', 'mountain quail', 5,
  'Mammalia', 'Sciuridae', 'medium mammal', 'douglas squirrel', 13,
  'Mammalia', 'Aplodontiidae', 'large mammal', 'mountain beaver', 3
)

# Copy out results from Ethier.
eth <- tribble(
  ~class, ~family, ~category, ~common, ~frequency,
  'Aves', 'Turdidae', 'medium bird', 'varied thrush', 39,
  'Aves', 'Corvidae', 'medium bird', "steller\'s jay", 38,
  'Aves', 'Picidae', 'medium bird', 'northern flicker', 34,
  'Aves', 'Alcidae', 'large bird', 'marbled murrelet', 15,
  'Aves', 'Columbidae', 'large bird', 'band-tailed pigeon', 5,
  'Aves', 'Corvidae', 'medium bird', 'gray jay', 4,
  'Aves', 'Hirundinidae', 'small bird', 'barn swallow', 2,
  'Aves', 'Strigidae', 'medium bird', 'northern pygmy owl', 1,
  'Aves', 'Turdidae', 'small bird', 'swainson\'s thrush', 1,
  'Aves', 'Tyrannidae', 'small bird', 'western wood-peewee', 1,
  'Mammalia', 'Sciuridae', 'medium mammal', 'red squirrel', 69,
  'Mammalia', 'Cricetidae', 'small mammal', 'vole', 1,
  'Mammalia', 'Unknown', 'small mammal', 'unknown', 1
)

# Ethier's results are WRONG!!!!! Must convert to percents.
eth <- eth %>% 
  mutate(total.count=sum(frequency), prop.count=frequency/total.count*100) %>% 
  select(-c(frequency, total.count))

# Add author information to each data frame.
blx <- blx %>% mutate(author='bloxton')
thr <- thr %>% mutate(author='thraikill')
#eth <- eth %>% mutate(author='ethier')

# Join the two frames together.
rv <- bind_rows(blx, thr)

# Add mass.
# rv <- left_join(rv, prey.list) %>% mutate_if(is.character, replace_na, 'Unknown') %>% 
#   left_join(mean.mass) %>% 
#   mutate(mass=coalesce(mass, average)) %>% 
#   select(author, class, family, order, genus, species, binomial, common, category, prop.count, mass)

# Calculate biomass.
#rv <- rv %>% mutate(biomass=prop.count*mass)

# Calculate counts from my own data.
# prop.count.avian.pr <- filter(diet.items, source != 'C') %>% 
#   mutate(total=n()) %>% filter(class == 'Aves') %>% 
#   mutate(count=n(), to.count=count/total*100) %>% 
#   distinct(to.count) %>% as.numeric()
# 
# prop.count.mammal.pr <- filter(diet.items, source != 'C') %>% 
#   mutate(total=n()) %>% filter(class == 'Mammalia') %>% 
#   mutate(count=n(), to.count=count/total*100) %>% 
#   distinct(to.count) %>% as.numeric()

# Add in new data.
mn <- tribble(
  ~author, ~class, ~proportion,
  'case', 'Aves', 60.55,
  'case', 'Mammalia', 39.45,
  'watson', 'Aves', 53,
  'watson', 'Mammalia', 47,
  'lewis', 'Aves', 83.65,
  'lewis', 'Mammalia', 16.35
)

# This is combined pellets-and-remains, for my own data as well as Watson's.

# Join with Bloxton & Thraikill.
pnw.class.summary <- rv %>% group_by(author, class) %>% 
  mutate(proportion=sum(prop.count)) %>% 
  distinct(author, class, proportion) %>% 
  bind_rows(mn) %>% 
  pivot_wider(names_from=class, values_from=proportion) %>% 
  arrange(Aves)

# For squirrels, also.
# prop.count.squirrel.pr <- filter(diet.items, source != 'C') %>% 
#   mutate(total=n()) %>% filter(group == 'squirrel') %>% 
#   mutate(count=n(), to.count=count/total*100) %>% 
#   distinct(to.count) %>% as.numeric()

mn.sq <- tribble(
  ~author, ~genus, ~prop.count,
  'case', 'Tamiasciurus', 24.22,
  'watson', 'Tamiasciurus', 15.1,
)

pnw.squirrel.summary <- rv %>% filter(common %in% c('douglas squirrel', 'red squirrel')) %>% 
  select(genus=common, prop.count, author) %>%
  bind_rows(mn.sq) %>% 
  arrange(prop.count)
```

Goshawk diet varies across its range in response to the regional presence and abundance of prey species (Drennan 2006). Tree squirrels are the same key prey identified on Vancouver Island (Ethier 1999) but differ from those in other regions of the Pacific Northwest, where the key prey is generally grouse (Watson et al. 1998, Thraikill et al. 2000, Bloxton 2002, Lewis et al. 2006). The key prey from my study area is more similar to those identified elsewhere in western North America, where the most important prey species is also mammalian and occasionally a sciurid. This unexpected result may be, in part, the product of differences in methodology. When the results from studies across temperate rainforest ecosystems are standardized (data from pooled pellets-and-remains or remains only, measured by counts), the difference between my study area and other regions in the Pacific Northwest is much less pronounced. However, the proportion of mammalian prey in the diet, particularly tree squirrels, remains markedly higher within British Columbia. This may be due to relatively high tree squirrel abundance within the south coast region, which is much greater than in other temperate rainforest ecosystems (Carey et al. 1995, Ransome & Sullivan 2003). No Pacific Northwest study has assessed goshawk diet and and absolute prey abundance simultaneously (but see Ethier 1999). Nonetheless, regional data hint at a pattern of higher dietary specialization in regions or forest types with higher tree squirrel abundance (see Figure \@ref(fig:squirrel-map)). Across the two ecological zones present in my study area I observed only minor variation in goshawk diet and no variation in the dominance of tree squirrels in the diet. If goshawks are more specialized on tree squirrels where tree squirrels are more abundant, this would indicate a slight difference in the prey community of these two zones but a similar abundance of tree squirrels. Goshawks in my study area appear to be a generalist predator opportunistically exploiting a locally abundant prey species.

```{r}
# Copy out approximate coordinate for all the study areas.
squirrel.coords <- tribble(
  ~lon, ~lat, ~author,
  #-125.621, 50.023, 'ethier',
  -122.019, 46.326, 'watson',
  -122.469, 49.588, 'case',
  -124.128, 47.999, 'carey95',
  -123.755, 47.464, 'bloxton',
  -123.842, 44.483, 'thraikill',
  -123.798, 42.685, 'carey92',
  -122.965, 49.335, 'ransome',
  -121.463, 48.614, 'carey93'
)

# Copy out squirrel densities (animals/ha)
squirrel.density <- tribble(
  ~type, ~value, ~author,
  'density', 0.2, 'carey92',
  'density', 0.1, 'carey95',
  'density', 0.5, 'carey93',
  'density', 1.1, 'ransome'
)

# Make the data set.
squirrel.map <- pnw.squirrel.summary %>% 
  mutate(genus='diet') %>% 
  rename(type=genus, value=prop.count) %>% 
  bind_rows(squirrel.density) %>% 
  left_join(squirrel.coords, by=c('author'))

# Scale it.
scale.map <- squirrel.map %>% pivot_wider(names_from=type, values_from=value) %>% 
  mutate(across(c(diet, density), ~scale(.) %>% as.vector)) %>% 
  pivot_longer(-c(author, lon, lat), names_to='type', values_to='value') %>% 
  drop_na() %>% left_join(squirrel.map, by=c('author', 'lon', 'lat', 'type'))

# Split out density and diet separately.
map.density <- filter(squirrel.map, type == 'density')
map.diet <- filter(squirrel.map, type == 'diet')

# Make a box for map size.
squirrel.box <- squirrel.coords %>% 
  st_as_sf(coords=c('lon', 'lat')) %>% 
  st_bbox()

# Bring in some base data.
n.america <- read_sf(dsn='../data/external/ne_10m_land.shp', layer='ne_10m_land')
rivers <- read_sf(dsn='../data/external/ne_10m_rivers_lake_centerlines.shp', 
                  layer='ne_10m_rivers_lake_centerlines')
```

```{r squirrel-map, out.width='100%', fig.cap='Comparison of goshawk dietary specialization and squirrel abundance in the Pacific Northwest. Size of symbol represents relative specialization or abundance. Goshawk diet estimated pellets-and-remains or remains only and measured using counts of items. Tree squirrel abundance estimated from number of individuals/ha. Adapted from Carey et al. (1992), Watson et al. (1998), Thrailkill et al. (2000), Bloxton (2002), Ransome and Sullivan (2003), and this study.'}
# Make a map.
ggplot() +
  geom_sf(data=n.america, fill='lightgrey') +
  geom_sf(data=rivers) +
  coord_sf(xlim=c(st_bbox(squirrel.box)[1] - 1.5, st_bbox(squirrel.box)[3] + 1.5), 
           ylim=c(st_bbox(squirrel.box)[2] - 0.25, st_bbox(squirrel.box)[4] + 0.25)) +
  geom_point(data=scale.map, aes(x=lon, y=lat, shape=type, size=value.x)) +
  geom_text_repel(data=map.density, aes(x=lon, y=lat, label=value),
                  direction='x',
                  nudge_x=5, # distance pushed
                  hjust=0, #horizontal justification to the left
                  segment.linetype='dashed') +
  geom_text_repel(data=map.diet, aes(x=lon, y=lat, label=round(value, digits=1)),
                  direction='x', hjust=1,
                  nudge_x=-5,
                  segment.linetype='dashed') +
  scale_shape_manual(values=c(15, 17), 
                     labels=c('squirrel density (squirrels/ha)', 'squirrel in diet (% items)')) +
  theme_void() +
  guides(size=FALSE) +
  theme(legend.title=element_blank(), panel.border=element_rect(color='black', fill=NA)) +
  annotation_north_arrow(location='bl', which_north='true', 
                         pad_y=unit(1, 'cm'), style=north_arrow_minimal) +
  annotation_scale(location='bl')
```

**Directions for future research**

<!--
  * Knowledge gaps
    * NOGO vs. NSPOW shows clear gaps in NOGO knowledge
    * Prey abundance - used to determine degree of selectivity, affects of prey abundance on pop dy
    * Space use - used to determine home range size, habitat selection
--> 

Comparing the northern spotted owl and the northern goshawk highlights significant knowledge gaps regarding goshawk biology. The controversy surrounding the spotted owl, combined with its position at the heart of a major management plan, has made it one of the most-studied birds in the world (Gutierrez et al. 2020). The northern goshawk, although likewise shrouded in controversy, has not received the same level of study. It is also more difficult to generalize research on the widespread, generalist northern goshawk than for the more restricted, relatively specialized spotted owl. As a result, there is a pressing need for detailed, local information on goshawks in the Pacific Northwest, particuarly in coastal British Columbia.

One knowledge gap related to the work I present here is how goshawks respond to variation in prey abundance. Dietary specialization such as goshawks in my study area exhibit may be the result of a preference for the main prey or simply an absence of other options. The hypotheses I propose to explain my results cannot be evaluated without differentiating between these two possibilities. Data on prey abundance is necessary to understand the mechanism of goshawk specialization and whether it is the result of high selectivity for tree squirrels, high abundance of tree squirrels, or some combination of both. This information could in turn be used to evaluate how changes in key prey abundance and total prey abundance influence goshawk demographic parameters, particularly productivity. Without this knowledge it is impossible to fully understand the factors limiting goshawk populations in coastal British Columbia.

A second knowledge gap is how the amount and type of habitat goshawks use for foraging varies with diet and prey abundance. Goshawk home range size varies across their geographic range and goshawks have been observed foraging in diverse habitat types, leading to persistent controversy over how and why goshawks select habitat, particularly foraging habitat (Greenwald et al. 2005, Reynolds et al. 2008). Research on spotted owls indicates that home range size is inversely related to prey abundance and foraging habitat selection varies with the identity of the main prey (Zabel 1995). Diet and prey abundance may likewise explain some of the variation in goshawk habitat use (Kenward 1982, Penteriani et al. 2013). If so, foraging habitat management could only be conducted with knowledge of goshawk diet and recommendations made for one population could be applied to others only with great caution, if at all. Filling this knowledge gap with information specific to coastal British Columbia is crucial and has been identified as a priority by goshawk managers (NGRT 2008, Parks Canada 2018).

<!--Parks Canada 2018: "Schedule of studies to identify critical habitat - continue to research critical habitat gaps - Knowledge of relationships between prey abundance, availability and diversity and foraging habitat characteristics will allow the refinement of specific foraging habitat targets for different habitat types." -->

<!-- 
  * Review of telemetry
    * Methods - tag info, capture technique
    * Results - home range size (males vs females), maximum range
-->

```{r message=FALSE, warning=FALSE, echo=FALSE}
# Read in table of nest coordinates.
telemetry.sites <- read_csv('../data/processed/telemetry_sites.csv')

# Twist wide so each id gets its own row.
captures <- select(telemetry.sites, m_tag, f_tag, year, date, site, nest) %>%
  pivot_longer(!c(year, date, site, nest), names_to='sex', values_to='id') %>%
  drop_na(id) %>%
  mutate(sex=case_when(
    sex == 'm_tag' ~ 'm',
    sex == 'f_tag' ~ 'f'
  )) %>% 
  distinct(nest, id, .keep_all=TRUE)

# Number of males and females.
n.males <- filter(captures, sex == 'm') %>% nrow()
n.females <- filter(captures, sex == 'f') %>% nrow()

# Bring in the clean telemetry data.
telemetry.data <- read_csv('../data/processed/20210420_clean_telemetry.csv', guess_max=15000)

# Define breeding season.
breeding.2018 <- interval(ymd(20180420), ymd(20180915))
breeding.2019 <- interval(ymd(20190420), ymd(20190915))
breeding.2020 <- interval(ymd(20200420), ymd(20200915))

# Select only points that fall within the breeding season.
breeding.telemetry <- telemetry.data %>% 
  filter(date %within% c(breeding.2018, breeding.2019, breeding.2020))

# Summary of days of telemetry data.
min.days <- breeding.telemetry %>% group_by(id) %>% 
  mutate(min=min(date), max=max(date), period=max-min) %>% 
  distinct(id, site, period) %>% ungroup() %>% slice_min(period) %>% 
  select(period) %>% as.numeric()

max.days <- breeding.telemetry %>% group_by(id) %>% 
  mutate(min=min(date), max=max(date), period=max-min) %>% 
  distinct(id, site, period) %>% ungroup() %>% slice_max(period) %>% 
  select(period) %>% as.numeric()

mean.days <- breeding.telemetry %>% group_by(id) %>% 
  mutate(min=min(date), max=max(date), period=max-min) %>% 
  distinct(id, site, period) %>% ungroup() %>% 
  summarize(mean=mean(period)) %>% as.numeric() %>% round(digits=2)

sd.days <- breeding.telemetry %>% group_by(id) %>% 
  mutate(min=min(date), max=max(date), period=max-min) %>% 
  distinct(id, site, period) %>% ungroup() %>% 
  summarize(sd=sd(period)) %>% as.numeric() %>% round(digits=2)

# Summary of location points.
mean.points <- breeding.telemetry %>% group_by(id, date) %>% 
  summarize(n=n()) %>% ungroup() %>% 
  summarize(mean=mean(n)) %>% as.numeric() %>% round(digits=2)

# Set of birds used to calculate home ranges.
home.range.set <- breeding.telemetry %>% group_by(id) %>% 
  mutate(min=min(date), max=max(date), n.points=n(), period=max-min) %>% 
  distinct(id, site, n.points, min, max, period) %>% 
  left_join(captures, by=c('id', 'site')) %>% 
  filter(period > 3)

# Number of birds in set.
n.males.hr <- filter(home.range.set, sex == 'm') %>% nrow()
n.females.hr <- filter(home.range.set, sex == 'f') %>% nrow()
```

To address this crucial knowledge gap, the British Columbia Ministry of Forests, Lands, Natural Resource Operations and Rural Development (FLNRORD) and I conducted a pilot study of foraging habitat use in coastal British Columbia. Between 2018-2020 we tagged `r n.males` male and `r n.females` female goshawks, with only one probable mortality over the course of the study. Captures were conducted during the early breeding season using a dho-gaza trap with a live great-horned owl (*Bubo virginianus*) as a lure (Bloom et al. 2007). Goshawks were fitted with 14-gram geolocator tags (Ecotone Telmetry brand, Harrier model) attached via backpack-style harness. Tags were programmed to take a location point every 15 minutes and to download data to a base station when within range (approximately 100 m from the base). Base stations were placed beneath or above nests and tags checked for proximity to the station every 15 minutes. From each tag we collected between 0 and `r max.days` days of breeding season (20 April - 15 September) data. Of the tags from which at least one day of data were retrieved, we collected between `r min.days` and `r max.days` (mean `r mean.days` $\pm$ `r sd.days` standard deviation) days of data, with an average of `r mean.points` locations per tag per day. <!--We excluded the individual with only three days of location data from most further analysis due to insufficient data, leaving us with a sample of `r n.males.hr` males and `r n.females.hr` females. -->

```{r echo=FALSE, warning=FALSE, message=FALSE}
# Make a spatial object.
telemetry.sf <- breeding.telemetry %>% 
  st_as_sf(coords=c('xcoord', 'ycoord'), remove=FALSE) %>%
  st_set_crs('+proj=utm +zone=10 +datum=WGS84 +units=m +no_defs')

# Calculate MCPs.
mcp <- telemetry.sf %>%
  select(id, geometry) %>%
  as_Spatial() %>%
  mcp.area(percent=c(50, 95), unin='m', unout='ha', plotit=FALSE) %>% 
  rownames_to_column(var='percent') %>% 
  pivot_longer(-percent, names_to='id', values_to='area') %>% 
  mutate(method='mcp')

# Calculate KDEs.
kde <- telemetry.sf %>%
  select(id, geometry) %>%
  as_Spatial() %>%
  kernelUD() %>% 
  kernel.area(percent=c(50, 95), unin='m', unout='ha') %>% 
  rownames_to_column(var='percent') %>% 
  pivot_longer(-percent, names_to='id', values_to='area') %>% 
  mutate(method='kde')

# And in the darkness bind them.
homerange <- bind_rows(mcp, kde)

# Also calculate maximum distance.
max.dist <- telemetry.sites %>% select(n.xcoord=x_coord, n.ycoord=y_coord, nest) %>% 
  right_join(breeding.telemetry, by=c('nest')) %>% 
  mutate(dist=sqrt((xcoord-n.xcoord)^2 + (ycoord-n.ycoord)^2)) %>% 
  group_by(id) %>% 
  arrange(desc(dist)) %>% 
  slice(1) %>% 
  select(id, dist, sex)

# Calculate number of points taken at nest.
nest.points <- breeding.telemetry %>% group_by(id) %>% 
  mutate(total.points=n(), base.points=sum(at.base), prop.at.base=base.points/total.points*100) %>% 
  distinct(id, sex, site, prop.at.base)
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Mean 95% MCP for all birds.
mean.95.mcp.hr <- homerange %>% 
  filter(method == 'mcp' & percent == 95) %>% 
  summarize(mean=mean(area)) %>% 
  as.numeric() %>% round(digits=2)

# Mean 95% KDE for all birds.
mean.95.kde.hr <- homerange %>% 
  filter(method == 'kde' & percent == 95) %>% 
  summarize(mean=mean(area)) %>% 
  as.numeric() %>% round(digits=2)

# Mean 95% MCP for males.
male.95.mcp.hr <- select(captures, id, sex) %>% 
  right_join(homerange, by=c('id')) %>% 
  filter(method == 'mcp' & percent == 95 & sex =='m') %>% 
  summarize(mean=mean(area)) %>% 
  as.numeric() %>% round(digits=2)

# Mean 95% KDE for males.
male.95.kde.hr <- select(captures, id, sex) %>% 
  right_join(homerange, by=c('id')) %>% 
  filter(method == 'kde' & percent == 95 & sex =='m') %>% 
  summarize(mean=mean(area)) %>% 
  as.numeric() %>% round(digits=2)

# Mean 95% MCP for females.
female.95.mcp.hr <- select(captures, id, sex) %>% 
  right_join(homerange, by=c('id')) %>% 
  filter(method == 'mcp' & percent == 95 & sex =='f') %>% 
  summarize(mean=mean(area)) %>% 
  as.numeric() %>% round(digits=2)

# Mean 95% KDE for females.
female.95.kde.hr <- select(captures, id, sex) %>% 
  right_join(homerange, by=c('id')) %>% 
  filter(method == 'kde' & percent == 95 & sex =='f') %>% 
  summarize(mean=mean(area)) %>% 
  as.numeric() %>% round(digits=2)

# Mean 50% MCP for males.
male.50.mcp.hr <- select(captures, id, sex) %>% 
  right_join(homerange, by=c('id')) %>% 
  filter(method == 'mcp' & percent == 50 & sex =='m') %>% 
  summarize(mean=mean(area)) %>% 
  as.numeric() %>% round(digits=2)

# Mean 50% MCP for females.
female.50.mcp.hr <- select(captures, id, sex) %>% 
  right_join(homerange, by=c('id')) %>% 
  filter(method == 'mcp' & percent == 50 & sex =='f') %>% 
  summarize(mean=mean(area)) %>% 
  as.numeric() %>% round(digits=2)

# Maximum distance for males.
male.max.dist <- max.dist %>% ungroup() %>% 
  filter(sex == 'm') %>% 
  summarize(mean=mean(dist)/1000) %>% 
  as.numeric() %>% round(digits=2)

# Maximum distance for females.
female.max.dist <- max.dist %>% ungroup() %>% 
  filter(sex == 'f') %>% 
  summarize(mean=mean(dist)/1000) %>% 
  as.numeric() %>% round(digits=2)

# Average points at nest for females.
female.nest.points <- nest.points %>% 
  filter(sex == 'f') %>% ungroup() %>% 
  summarize(mean=mean(prop.at.base)) %>% 
  as.numeric() %>% round(digits=2)

# Average points at nest for males.
male.nest.points <- nest.points %>% 
  filter(sex == 'm') %>% ungroup() %>% 
  summarize(mean=mean(prop.at.base)) %>% 
  as.numeric() %>% round(digits=2)

# SD points at nest for females.
female.nest.sd <- nest.points %>% 
  filter(sex == 'f') %>% ungroup() %>% 
  summarize(sd=sd(prop.at.base)) %>% 
  as.numeric() %>% round(digits=2)

# SD points at nest for males.
male.nest.sd <- nest.points %>% 
  filter(sex == 'm') %>% ungroup() %>% 
  summarize(sd=sd(prop.at.base)) %>% 
  as.numeric() %>% round(digits=2)
```

The mean breeding season home range was `r mean.95.mcp.hr` ha (95% minimum convex polygon) but I observed a large difference in home range size between males and females. Mean male home range was `r male.95.mcp.hr` ha whereas mean female home range was much smaller, `r female.95.mcp.hr` ha. Core-use areas (50% minimum convex polygon) were more similar, but still larger for males at `r male.50.mcp.hr` ha than for females at `r female.50.mcp.hr` ha. The smaller female home range was due to the amount of time females spent on the nest. An average of `r female.nest.points`% ($\pm$ `r female.nest.sd`) of female location points were taken within range of the base station (< 100 m from the nest), while only `r male.nest.points`% ($\pm$ `r male.nest.sd`) of male points were taken at the nest. Males also traveled further from the nest. The mean furthest distance from the nest a male traveled was `r male.max.dist` km, while females on average only ventured `r female.max.dist` km from the nest. As the breeding season progressed, the length of time spent away from the nest increased noticeably for males but less so for females. <!-- needs recurse -->

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Make the summary table.
summary.table <- breeding.telemetry %>% group_by(id) %>% 
  mutate(min=min(date), max=max(date), n.points=n(), period=max-min) %>% 
  distinct(id, site, n.points, min, max, period) %>% 
  left_join(max.dist, by=c('id')) %>% 
  left_join(homerange, by=c('id')) %>% 
  left_join(nest.points, by=c('id', 'site', 'sex')) %>% 
  pivot_wider(names_from=c(method, percent), names_sep='.', values_from=area) %>% 
  arrange(sex, site) %>%
  rownames_to_column(var='order') %>% mutate(order=as.numeric(order)) %>% 
  mutate(order=case_when(
    sex == 'f' ~ order - 1,
    sex == 'm' ~ order,
    TRUE ~ order
  ),
  period=as.numeric(period)) %>% 
  mutate(., across(where(is.numeric), round, digits=2))

# Find where total rows should be inserted.
table.breaks <- summary.table %>% group_by(sex) %>% 
  summarize(max=max(order), .groups='drop') %>% ungroup() %>% 
  select(max) %>% as.vector()

# Summarize female data.
female.summary <- summary.table %>% filter(sex == 'f') %>% ungroup() %>%  
  summarise(., across(where(is.numeric), mean)) %>% 
  mutate(order=table.breaks$max[1] + 1, id='AVERAGE', site=NA, min=NA, max=NA, n.points=NA, sex='f')

# Summarize male data.
male.summary <- summary.table %>% filter(sex == 'm') %>% ungroup() %>%  
  summarise(., across(where(is.numeric), mean)) %>% 
  mutate(order=table.breaks$max[2] + 1, id='AVERAGE', site=NA, min=NA, max=NA, n.points=NA, sex='m')

bind_rows(summary.table, female.summary, male.summary) %>% 
  arrange(order) %>% column_to_rownames(var='order') %>% 
  select(-sex) %>% relocate(n.points, .after=period) %>% 
  flextable() %>% 
  colformat_num(na_str='-', digits=0) %>% 
  add_header_row(values=c(' ', 'Days of data collected', 'Relation to nest', 'Home range (ha)'), 
                 colwidths=c(2, 3, 3, 4)) %>% 
  set_header_labels(id='ID', site='Site', min='first', max='last', n.points='N. points',
                    period='N. days', dist='Max. distance from nest (m)', 
                    prop.at.base='% points taken at nest',
                    mcp.50='50% MCP', mcp.95='95% MCP', kde.50='50% KDE', kde.95='95% KDE')
```

```{r echo=FALSE, message=FALSE, warning=FALSE, eval=FALSE}
select(captures, id, sex) %>% right_join(homerange, by=c('id')) %>% 
  group_by(method, sex, percent) %>% 
  summarize(n=n(), mean.area=mean(area)) %>% 
  pivot_wider(names_from=method, values_from=mean.area) %>% 
  ungroup() %>% #select(percent, mcp, kde) %>% 
  flextable() %>% 
  add_header_row(values=c(' ', 'area (ha)'), colwidths=c(3, 2))
```

<!--By the end of the study period, one goshawk was a probable mortality, one's fate was unknown, five were likely resident near their capture sites but had not been conclusively relocated since the season they were tagged, and six were still extant.-->

We encountered several challenges over the course of the three-year pilot. The dho-gaza method is biased toward the capture of females and resulted in a larger number of females than males in our sample. Because males provide most of the food during the breeding season, this provided us with relatively little foraging location data. Further, the geolocator tags only download data on a 15-minute cycle and when within 100 m of the base station. Males spend little time near the nest, save briefly while making prey deliveries, so we obtained only infrequent data downloads from males. Overall, our methods favored the collection of location data from females, which offered relatively little insight into foraging habitat use. We also struggled to relocate tagged birds in subsequent years and so failed to obtain location data for multiple breeding seasons or winter movement. It is difficult to determine the fate of these missing birds. Several were tentatively resighted but could not be definitely relocated, and it is unclear if they nested the year after being tagged. We cannot rule out the possibility that tags negatively impacted these birds' ability to initiate. However, we observed only one probable mortality and no nest failures or abandonment among tagged birds, which suggests tag effects were not large. The location data we were able to retrieve, although less than anticipated, was of much higher resolution than VHF data and much higher accuracy than satellite data. Despite setbacks, this technology may therefore be uniquely well-suited to studies of foraging habitat selection.

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Make a map.

# Calculate 50% and 95% MCPs for female.
rlk.f.mcp.95 <- telemetry.sf %>%
  filter(id == 'HAR02') %>% 
  select(id, geometry) %>%
  as_Spatial() %>%
  mcp(percent=95, unin='m', unout='ha') %>% 
  st_as_sf()

# 50% MCP
rlk.f.mcp.50 <- telemetry.sf %>%
  filter(id == 'HAR02') %>% 
  select(id, geometry) %>%
  as_Spatial() %>%
  mcp(percent=50, unin='m', unout='ha') %>% 
  st_as_sf()

# Calculate 50% and 95% MCPs for male.
rlk.m.mcp.95 <- telemetry.sf %>%
  filter(id == 'HAR04') %>% 
  select(id, geometry) %>%
  as_Spatial() %>%
  mcp(percent=95, unin='m', unout='ha') %>% 
  st_as_sf()

rlk.m.mcp.50 <- telemetry.sf %>%
  filter(id == 'HAR04') %>% 
  select(id, geometry) %>%
  as_Spatial() %>%
  mcp(percent=50, unin='m', unout='ha') %>% 
  st_as_sf()

# Pull out points for each.
rlk.m.points <- telemetry.sf %>% 
  filter(id == 'HAR04')

rlk.f.points <- telemetry.sf %>% 
  filter(id == 'HAR02')

# Pull out points for all.
rlk.points <- telemetry.sf %>% 
  filter(site == 'RLK')

# Pull out the nests.
rlk.nests <- telemetry.sites %>% 
  filter(site == 'RLK') %>% 
  st_as_sf(coords=c('x_coord', 'y_coord'), remove=FALSE) %>%
  st_set_crs('+proj=utm +zone=10 +datum=WGS84 +units=m +no_defs')

# Calculate a centroid.
rlk.centroid <- telemetry.sites %>% 
  filter(site == 'RLK') %>% 
  summarize(mean.x=mean(x_coord), mean.y=mean(y_coord)) %>% 
  st_as_sf(coords=c('mean.x', 'mean.y'), remove=FALSE) %>%
  st_set_crs('+proj=utm +zone=10 +datum=WGS84 +units=m +no_defs')

# Calculate a buffer around each nest.
area <- 200

radius <- sqrt(area*10000/pi)

rlk.buffer <- st_buffer(rlk.centroid, radius)
rlk.hr <- st_buffer(rlk.centroid, 3450)
```

```{r telemetry-map, fig.cap='Location data from female (a) and male (b) goshawks'}
ggplot() +
  geom_sf(data=rlk.f.points, alpha=0.25) +
  geom_sf(data=rlk.f.mcp.50, fill='white') +
  geom_sf(data=rlk.f.mcp.95, fill=NA) +
  geom_sf(data=rlk.buffer, fill=NA, linetype='dashed') +
  geom_sf(data=rlk.hr, fill=NA, linetype='dashed') +
  geom_sf(data=rlk.centroid, color='red', size=2) +
  annotation_scale(location='br') +
  theme_void() +
  coord_sf(xlim=c(st_bbox(rlk.points)[1] - 0.5, st_bbox(rlk.points)[3] + 0.5), 
           ylim=c(st_bbox(rlk.points)[2] - 0.25, st_bbox(rlk.points)[4] + 0.25)) +
  theme(panel.border=element_rect(color='black', fill=NA),
        legend.position='bottom') +
  labs(title='a')

ggplot() +
  geom_sf(data=rlk.m.mcp.50, fill=NA) +
  geom_sf(data=rlk.m.mcp.95, fill=NA) +
  geom_sf(data=rlk.buffer, fill=NA, linetype='dashed') +
  geom_sf(data=rlk.hr, fill=NA, linetype='dashed') +
  geom_sf(data=rlk.m.points, alpha=0.25) +
  geom_sf(data=rlk.nests, color='red', size=2) +
  annotation_scale(location='br') +
  theme_void() +
  coord_sf(xlim=c(st_bbox(rlk.points)[1] - 0.5, st_bbox(rlk.points)[3] + 0.5), 
         ylim=c(st_bbox(rlk.points)[2] - 0.25, st_bbox(rlk.points)[4] + 0.25)) +
  theme(panel.border=element_rect(color='black', fill=NA),
        legend.position='bottom') +
  labs(title='b')
```

My results, though preliminary, show several interesting patterns and offer promising suggestions for management. The high fidelity of females to the nest area confirms the importance of protected buffers around nests to prevent the disturbance of critical habitat. However, the large distances traveled by males and the irregular shape of their home ranges shows that circular buffers based on estimates of mean home range size are not good approximations of true space use. Home ranges were rarely circular, but rather shaped by geographic constraints such as coastlines or mountains peaks. Within home ranges, kernel density estimates show habitat use is not even across a male's territory. Instead, there are clusters of activity around areas which are likely high-quality foraging habitat. These areas usually fall outside the nest area and are not be captured by current nesting habitat protections. Goshawks readily crossed narrow barriers such as roads, rivers, and powerline cuts, and sometimes appeared to use these features for travel or foraging. On the other hand, goshawks seemed to generally route around larger barriers such as lakes or clearcuts. This suggests managed areas around nests should be tailored to the amount and configuration of suitable foraging habitat present in the surrounding landscape, and that any timber harvest within the foraging area should prioritize the preservation of connectivity between patches of high-quality foraging habitat. However, a great deal of additional research is needed to confirm and refine these suggestions, particularly regarding the characteristics of suitable and high-quality foraging habitat.

<!-- Carey et al. 1992 (re: maximum distance): "The old forest set aside must be within some area the owl can reasonably traverse. Our range lengths suggest that 5 km is a reasonable distance for owls to traverse on an annual basis. A circle of 2.5 km radius is 2000 ha. Such a circle could be used as a template. If > 950 ha of old forest were present within the 2000 ha circle, there is a reasonably high expectation that the area could support a pair of owls for 1 year." -->

**Management implications**

Climate change is one of the largest and most difficult to anticipate threats facing biodiversity (Bellard et al. 2012). The current understanding of how climate change will impact goshawks in coastal British Columbia is largely speculative (Parks Canada 2018). Although I found no evidence of a link between the degree of dietary specialization and goshawk productivity, the level of specialization I observed and findings from other studies suggest goshawks in my study area are highly dependent on tree squirrels for successful reproduction. Notably, Ethier (1999) found goshawk productivity on Vancouver Island to be significantly correlated with red squirrel abundance. Tree squirrels are themselves highly dependent on conifer seeds and their abundance varies with the size of the annual cone crop. (Smith 1970)<!-- check this and find better/more specific -->. Seed production is cyclical, but may be increased by higher temperatures and inhibited by drought (Boucher et al. 2020). As the climate warms and drought risk rises, climate-driven changes in seed production may cascade upward through prey abundance to impact goshawk populations in coastal British Columbia. Significantly more work is needed to understand the relationship between prey abundance, the buffering ability of alternate prey, and goshawk productivity, but my results provide an important starting point for incorporating climate change into goshawk management.

<!-- Parks Canada 2018: "The threats caused by climate change within coastal forests are unknown and are difficult to mitigate through this recovery strategy. Climate change may affect Northern Goshawk
populations positively or negatively." -->

<!-- Ethier 1999: "Correlation of goshawk abundance and productivity with red squirrel abundance revealed strong positive correlations." -->

Wildlife management, whether rooted in a single-species or ecosystem-based approach, generally centers around species, subspecies, or unique populations at risk of extinction (cite). My thesis examines a single population of goshawks on the south coast of British Columbia which is currently classified as part of the subspecies *laingi*. The precise range of *laingi* has always been unclear, but recent genetic evidence calls into question the boundaries currently used in goshawk management. Geraldes et al. (2019) identified the Haida Gwaii population as highly unique and distinct from populations on Vancouver Island and the mainland coast. Their findings indicate the distribution of *laingi* may be restricted to Haida Gwaii and challenge the conservation status of goshawk populations elsewhere within the putative range of *laingi*. My results suggest a strong ecological similarity between goshawks within my study area, Vancouver Island, and Haida Gwaii, where goshawks are also highly dependent on tree squirrels (COSEWIC 2013). The diet of these populations stands in contrast to others in temperate rainforest ecosystems which are currently considered *laingi*, such as the Olympic Peninsula or Southeast Alaska, where goshawks consume mainly birds. It is further different from populations elsewhere in western North America, where goshawks also consume primarily mammals but do not specialize on tree squirrels. Goshawks in coastal British Columbia may therefore be ecologically similar to each other, and ecologically distinct from other populations, even if they are genetically dissimilar. This does not indicate a single, uniform management strategy is appropriate for all goshawks in the region: red squirrels, for instance, are an introduced species on Haida Gwaii. It does, however, demonstrate that management plans benefit from the type of fine-scale ecological information my thesis provides. It further suggests these goshawks may deserve special management consideration as a population unique to the temperate rainforests of coastal British Columbia.