---
title: "Chapter Two: Methods 0.1"
output: html_notebook
always_allow_html: true
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE)
options(knitr.kable.NA = '-')
```

# Methods

```{r libraries, message=FALSE, warning=FALSE}
# Import conflict settings.
source('../src/conflicted.R')

# Load some libraries.
library(tidyverse)
library(knitr)
library(kableExtra)
library(flextable)
library(sf)
library(broom)
library(AICcmodavg)
```

```{r load-rasters, eval=FALSE, message=FALSE}
## For making rasters of the fHSI data, see 'rasterizing_hsi.R'.
## For making rasters of the VRI data, see 'rasterizing_vri.R'.

library(raster)

# Load all the rasters.
r.bec <- raster('../data/external/bec_raster_full.tif')
r.landcover <- raster('../data/processed/vri_sc_all-sites.tif')
r.older <- raster('../data/processed/older_sc_100.tif')
r.gaps <- raster('../data/processed/gaps_sc_100.tif')
r.canopy <- raster('../data/processed/canopy_sc_100.tif')
r.hsi <- raster('../data/processed/foraging_sc.tif')
r.suitable <- raster('../data/processed/foraging_sc_suitable.tif')

# Define levels for each raster.
bec.levels <- levels(r.bec) %>% data.frame()

landcover.levels <- data.frame(ID=0:11, class.name=
                                 c('undefined', 'rock', 'ocean', 'freshwater',
                                   'alpine', 'wetland',
                                   'shrub', 'deciduous', 'regen',
                                   'young', 'mature', 'old'))

older.levels <- data.frame(ID=0:1, class.name=
                             c('undefined', 'older'))

gap.levels <- data.frame(ID=0:1, class.name=
                           c('undefined', 'gap'))

canopy.levels <- data.frame(ID=c(-1, 0, 50, 75), 
                            class.name=c('undefined', 'none', 'moderate', 'high'))

hsi.levels <- data.frame(ID=c(-10, -2, -1, 0, 1, 2, 3), 
                         class.name=c('ocean', 'freshwater', 'river', 
                                      'nil', 'low', 'moderate', 'high'))

suitable.levels <- data.frame(ID=c(0, 4), 
                              class.name=c('unsuitable', 'suitable'))

# Add levels to each raster.
levels(r.landcover) <- landcover.levels
levels(r.older) <- older.levels
levels(r.gaps) <- gap.levels
levels(r.canopy) <- canopy.levels
levels(r.hsi) <- hsi.levels
levels(r.suitable) <- suitable.levels

# Finally, assign a CRS.
lapply(c(r.bec, r.landcover, r.older, r.gaps, r.canopy, r.hsi, r.suitable), 
       function(x) {crs(x) <- CRS('+proj=utm +zone=10 +datum=NAD83 +units=m +no_defs')})

# Lapply doesn't work for several, so add manually.
crs(r.bec) <- CRS('+proj=utm +zone=10 +datum=NAD83 +units=m +no_defs')
crs(r.hsi) <- CRS('+proj=utm +zone=10 +datum=NAD83 +units=m +no_defs')
crs(r.suitable) <- CRS('+proj=utm +zone=10 +datum=NAD83 +units=m +no_defs')
```

```{r calc-metric, eval=FALSE, message=FALSE, warning=FALSE}
#source('../src/load_rasters.R')
library(landscapemetrics)

### Mature forest.

# Make a list of landcover metrics to calculate.
mature.metrics <- c('lsm_c_pland', 'lsm_l_ed')

# Make a function to do the calculations and formatting.
calc.mature.metrics <- function(x) {
  sample_lsm(r.older, y=sites.sf, size=x, plot_id=site.names, shape='circle', 
             what=mature.metrics) %>% 
    left_join(older.levels, by=c('class'='ID')) %>% 
    mutate(class.name=ifelse(is.na(class.name), metric, class.name)) %>% 
    select(-class, -metric, -level) %>%  
    pivot_wider(names_from=class.name, values_from=value) %>% 
    mutate(radius=x)%>% 
    rename(mature.inside=percentage_inside)
}

# Run the function for each sample size.
mature.landscape.metrics <- map_df(landscape$radius, calc.mature.metrics)

mature.landscape.metrics <- mature.landscape.metrics %>% 
  replace_na(list(undefined=0, older=0)) %>% 
  select(radius, nest=plot_id, mature.inside,
         mature.edge.density=ed, 
         proportion.mature=older, proportion.undefined=undefined)

mature.landscape.metrics <- select(landscape, radius, size) %>%
  right_join(mature.landscape.metrics)

### Gaps

# Make a list of gap metrics to calculate.
gap.metrics <- c('lsm_l_ed')

# Make a function to do the calculations and formatting.
calc.gap.metrics <- function(x) {
  sample_lsm(r.gaps, y=sites.sf, size=x, plot_id=site.names, shape='circle', 
             what=gap.metrics) %>% 
    left_join(gap.levels, by=c('class'='ID')) %>% 
    mutate(class.name=ifelse(is.na(class.name), metric, class.name)) %>% 
    select(-class, -metric, -level) %>%  
    pivot_wider(names_from=class.name, values_from=value) %>% 
    mutate(radius=x) %>% 
    rename(gap.inside=percentage_inside)
}

# Run the function for each sample size.
gap.landscape.metrics <- map_df(landscape$radius, calc.gap.metrics)

# Tidy things up.
gap.landscape.metrics <- gap.landscape.metrics %>% 
  select(nest=plot_id, gap.edge.density=ed, radius, gap.inside)

gap.landscape.metrics <- select(landscape, radius, size) %>% right_join(gap.landscape.metrics)

### Canopy cover.

# Make a list of canopy metrics to calculate.
canopy.metrics <- c('lsm_c_pland', 'lsm_c_ed')

# Make a function to do the calculations and formatting.
calc.canopy.metrics <- function(x) {
  sample_lsm(r.canopy, y=sites.sf, size=x, plot_id=site.names, shape='circle', 
             what=canopy.metrics) %>% 
  left_join(canopy.levels, by=c('class'='ID')) %>% 
  mutate(class.name=ifelse(is.na(class.name), metric, class.name)) %>% 
  select(-class, -level) %>%  
  pivot_wider(names_from=c(class.name, metric), values_from=value, names_sep='.') %>% 
  mutate(radius=x) %>% 
  rename(canopy.inside=percentage_inside)
}

# Run the function for each sample size.
canopy.landscape.metrics <- map_df(landscape$radius, calc.canopy.metrics)

# Do some cleanup: fill NAs with zeros and rename columns, calculate mature forest.
canopy.landscape.metrics <- canopy.landscape.metrics %>% 
  replace_na(list(none.plant=0, moderate.pland=0, high.pland=0)) %>% 
  select(radius, nest=plot_id, canopy.inside,
         proportion.canopy.high=high.pland,
         canopy.high.edge.density=high.pland)

canopy.landscape.metrics <- select(landscape, radius, size) %>% 
  right_join(canopy.landscape.metrics, by=c('radius'))

### HSI

# Make a list of metrics to calculate.
hsi.metrics <- c('lsm_l_ed')

# Make a function to do the calculations and formatting.
calc.hsi.metrics <- function(x) {
  sample_lsm(r.hsi, y=sites.sf, size=x, plot_id=site.names, shape='circle', 
             what=hsi.metrics) %>% 
    left_join(hsi.levels, by=c('class'='ID')) %>% 
    mutate(class.name=ifelse(is.na(class.name), metric, class.name)) %>% 
    select(-class, -metric, -level) %>%  
    pivot_wider(names_from=class.name, values_from=value) %>% 
    mutate(radius=x) %>% 
    rename(hsi.inside=percentage_inside)
}

# Run the function for each sample size.
hsi.landscape.metrics <- map_df(landscape$radius, calc.hsi.metrics)

# Do some cleanup
hsi.landscape.metrics <- hsi.landscape.metrics %>% #replace(is.na(.), 0) %>% 
  select(radius, hsi.inside, nest=plot_id, hsi.edge.density=ed)

hsi.landscape.metrics <- select(landscape, radius, size) %>% 
  right_join(hsi.landscape.metrics, by=c('radius'))

### Suitable habitat.

# Make a list of metrics to calculate.
suitable.metrics <- c('lsm_c_pland', 'lsm_l_ed')

# Make a function to do the calculations and formatting.
calc.suitable.metrics <- function(x) {
  sample_lsm(r.suitable, y=sites.sf, size=x, plot_id=site.names, shape='circle', 
             what=suitable.metrics) %>% 
    left_join(hsi.levels, by=c('class'='ID')) %>% 
    mutate(class.name=ifelse(is.na(class.name), metric, class.name)) %>% 
    select(-class, -metric, -level) %>%  
    pivot_wider(names_from=class.name, values_from=value) %>% 
    mutate(radius=x) %>% 
    rename(suitable.inside=percentage_inside)
}

# Run the function for each sample size.
suitable.landscape.metrics <- map_df(landscape$radius, calc.suitable.metrics)

# Do some cleanup
suitable.landscape.metrics <- suitable.landscape.metrics %>% # replace(is.na(.), 0) %>% 
  select(radius, suitable.inside, nest=plot_id, suitable.edge.density=ed, proportion.suitable=pland)

suitable.landscape.metrics <- select(landscape, radius, size) %>% 
  right_join(suitable.landscape.metrics, by=c('radius'))

### Wrap-up.

landscape.data <- full_join(mature.landscape.metrics, suitable.landscape.metrics, 
                            by=c('nest', 'radius', 'size')) %>% 
  full_join(hsi.landscape.metrics, by=c('nest', 'radius', 'size')) %>% 
  full_join(gap.landscape.metrics, by=c('nest', 'radius', 'size')) %>% 
  full_join(canopy.landscape.metrics, by=c('nest', 'radius', 'size'))

write_csv(landscape.data, '../data/processed/landscape_metrics_plain.csv')
```

## Study Area and Species

In North America, the northern goshawk ranges from boreal forests of the Yukon south to high-elevation forests of Arizona and New Mexico. Two subspecies are recognized: the widespread *atricapillus* and the limited *laingi* (Squires et al. 2020). The *laingi* subspecies was first described on the Haida Gwaii archipelago in British Columbia and is smaller and darker than the *atricapillus* subspecies found elsewhere on the continent (Taverner 1940). The range of this subspecies is limited to the west coast of North America from southeast Alaska through mainland British Columbia and Vancouver Island, possibly as far south as Washington’s Olympic Peninsula (COSEWIC 2013). *A. g. laingi* is considered a species at risk in British Columbia by both the federal and provincial governments due to significant habitat loss from industrial timber harvest (Team 2008, COSEWIC 2013).

Several models have been proposed to describe the components of a goshawk territory and capture the hierarchical nature of goshawk habitat selection (reviewed in Squires & Kennedy 2005, Andersen et al. 2006). At the largest scale, the home range or foraging area includes the total area used by the male and female during the breeding season. Within the home range, the breeding area is the core-use area for adults and is used by fledglings across multiple years prior to dispersal. At the smallest scale around each nest, the post-fledging area is used by fledglings within a single year after fledging but prior to dispersal (*sensu* McClaren et al. 2015). Goshawks often build multiple nests within the breeding area, and each nest has an associated post-fledging area.

We studied goshawks in southwestern British Columbia, a region characterized by rugged mountains interspersed with coastal fjords and low-lying valleys. The maritime climate supports temperate rainforest dominated by Douglas-fir (*Pseudotsuga menziesii*), western redcedar (*Thuja plicata*), and western hemlock (*Tsuga heterophylla*) (Meidinger and Pojar 1991). The goshawk population in southwestern British Columbia is currently classified as *A. g. laingi*, though new genetic evidence may lead to future reclassification (Geraldes et al. 2018). 

## Goshawk territories

```{r occupancy-data, message=FALSE, warning=FALSE}
# Bring in occupancy data.
occupancy.history <- read_csv('../data/processed/occupancy_sc.csv')
## In this dataset: 
# 0 = not visited
# 1 = visited, no detection
# 2 = visited, with detection
# 3 = visited, breeding

# Remove problem child Turbid Creek. This site has a known error with the GIS data.
occupancy.history <- filter(occupancy.history, site != 'TCR')

# Remove any sites that have never been visited (status in all years = 0).
occupancy.history <- occupancy.history %>% 
  filter_at(vars(starts_with('status')), any_vars(. > 0))

# Twist occupancy data long.
occupancy.data <- occupancy.history %>% 
  pivot_longer(-c(site, name), names_to='year', values_to='status') %>% 
  filter(status > 0) %>%
  group_by(site, status) %>% 
  add_tally() %>% 
  distinct(site, status, .keep_all=TRUE) %>% 
  select(-year) %>% 
  pivot_wider(names_from=status, values_from=n, values_fill=0) %>% 
  ungroup() %>% rowwise(site, name) %>% 
  mutate(years.surveyed=sum(c(`3`, `2`, `1`)),
         years.detect=sum(c(`3`, `2`)),
         years.no.detect=years.surveyed-years.detect) %>%
  select(site, name, years.surveyed, years.detect, years.no.detect)

# Bring in calculated landscape data.
landscape.data <- read_csv('../data/processed/landscape_metrics_plain.csv')

# Join to occupancy data.
occupancy.data <- occupancy.data %>% left_join(landscape.data, by=c('site'='nest'))

# Filter out everything with insufficient GIS coverage (less than 90%).
occupancy.data <- occupancy.data %>% group_by(site) %>% 
  filter(across(contains('inside'), ~all(.>=90)))
```

Goshawk territories were located as part of long-term population monitoring conducted by the British Columbia Ministry of Forests, Lands, Natural Resource Operations and Rural Development (FLNRORD). Nest locations were collected by from 2013-2020 by FLNRO surveyors, timber industry professionals, and members of the public and compiled in a database of goshawk territories within the study area. From this dataset we removed all territories which had never been visited by a FLRNORD surveyor, or which lacked sufficient GIS data coverage (see below), producing a set of `r distinct(occupancy.data, site) %>% nrow()` territories retained for analysis. Territories were not surveyed every year, and most territories had a single survey visit for each year it was surveyed. We defined a territory as "active" if a goshawk was detected during a survey or definitive signs of breeding activity were observed, and "not active" if no goshawks were detected. A single survey visit has a low (%, source) probability of detecting breeding activity and the probability of detection increases with survey effort. We made no attempt to correct for survey effort and acknowledge that our data likely underestimate goshawk activity.

## Landscape metrics

```{r nest-centroids, message=FALSE, warning=FALSE}
# Bring in nest data.
nests <- read_csv('../data/processed/sc_nests.csv')

# Calculate a centroid for each site.
centroids <- nests %>% group_by(site) %>% 
  mutate(mean.x=mean(xcoord), mean.y=mean(ycoord)) %>% 
  distinct(site, name, mean.x, mean.y)

sites <- centroids %>% rename(xcoord=mean.x, ycoord=mean.y)

# Make it a spatial object for later.
sites.sf <- sites %>% 
  st_as_sf(coords=c('xcoord', 'ycoord')) %>% 
  st_set_crs('+proj=utm +zone=10 +datum=WGS84 +units=m +no_defs') %>% 
  st_as_sf()

# Also make a list of site names for later.
site.names <- sites.sf$site
```

```{r spatial-scales}
# Set up the smaller landscape scales.
landscape <- data.frame(
  size=c('PFA', 'breeding area', 'home range'),
  area=c(60, 200, 3800)
)

# Calculate the maxiumum range.
max.dist <- c(7773, 8023, 7977, 4413) %>% mean()
## These numbers are from four males goshawks tagged in the South Coast in 2019. 
## Data are from 20 April to 15 September.

# Convert to area.
max.area <- (pi*max.dist^2)/10000

# Add to landscapes.
landscape <- data.frame(
  size='maximum range',
  area=max.area
) %>% 
  bind_rows(landscape) %>% arrange(area)

# Convert area in hectares to radii in meters.
landscape <- landscape %>% mutate(radius=sqrt(area*10000/pi))
```

We quantified habitat composition and configuration at four scales (60 ha, 200 ha, 3800 ha, and 15600 ha) around the center of each goshawk territory. Territory centers were calculated by taking the mean location of all nests within each territory. The three smallest scales (60, 200, and 3800 ha) were defined using estimates of the post-fledging area (McClaren 2005), breeding area (FLNRO 2013 or whatever, and McClaren 2005 again?), and breeding season home range (McClaren et al. 2015 works but actually McClaren... 2005?), respectively. Because the space-use of goshawks in southwestern British Columbia is largely unknown, these estimates come from Vancouver Island, British Columbia, which is ecologically similar to our study area. The largest scale (15600 ha) is based on the mean maximum distance traveled from the nest during the breeding seasons by four geolocator-tagged male goshawks during the breeding season (unpublished data). This number attempts to capture the largest extent available for goshawk foraging activities.

We characterized goshawk habitat using data from two sources. A habitat suitability model (USFWS 1981) for goshawk foraging habitat was developed for the coastal goshawk population to aid in management planning and action (Mahon 2008). From this model we extracted four indices of suitability and further derived two simplified categories of suitability. Indices of suitability were high suitability (3), moderate suitability (2), low suitability (1), and unsuitable (0). The simplified categories were suitable (2-3) and unsuitable (0-1). From vegetation mapping data (cite) we defined two landcover classes and four canopy cover classes. Landcover classes were mature forest (> 70 years old) and openings (shrubby areas, regenerating burns, and very young forest < 5 years old). Canopy cover classes were low cover (< 50%), moderate cover (50-75%), and high cover (> 75%). 

We then calculated three composition metrics and five fragmentation metrics at each spatial scale. Composition metrics were the proportion of suitable habitat, mature forest, and high canopy cover. Fragmentation metrics were the edge density of all habitat suitability classes, suitable habitat,  openings, mature forest, and all canopy cover classes.

```{r correlation, eval=FALSE}
# Correleogram for the suitability variables.
occupancy.data %>% ungroup() %>% 
  select(proportion.suitable, suitable.edge.density, size) %>% 
  ggpairs(columns=1:2, ggplot2::aes(color=size))

# Correleogram for the forest age variables.
occupancy.data %>% ungroup() %>% 
  select(proportion.mature, mature.edge.density, size) %>% 
  ggpairs(columns=1:2, ggplot2::aes(color=size))
```

We developed a set of nine generalized linear models to test hypotheses regarding the influence of forest characteristics at different scales on the probability of goshawk activity. We predicted the probability of goshawk use would increase with greater amounts of preferred habitat, as measured by the habitat suitability index or forest age, and decrease with greater amounts of fragmentation. We also predicted this trend would be stronger at small scales closer to the nest. We fit models using the number of years a territory was active as the response variable, weighted by the number of years a territory was surveyed, with the binomial family and a logit link. We ranked models using Aikake's Information Criterion and retained models with a delta-AIC < 2 for each scale.

```{r models}
# Proportion suitable.
proportion.suitable.model <- function(df) {
  glm(cbind(years.detect, years.no.detect) ~ proportion.suitable, data=df, family=binomial(logit))
}

# Proportion suitable + suitable edge density.
suitable.plus.suitable.edge.model <- function(df) {
  glm(cbind(years.detect, years.no.detect) ~ proportion.suitable + suitable.edge.density, data=df, family=binomial(logit))
}

# Proportion suitable + HSI edge density.
suitable.plus.hsi.edge.model <- function(df) {
  glm(cbind(years.detect, years.no.detect) ~ proportion.suitable + hsi.edge.density, data=df, family=binomial(logit))
}

# Proportion mature forest.
proportion.mature.model <- function(df) {
  glm(cbind(years.detect, years.no.detect) ~ proportion.mature, data=df, family=binomial(logit))
}

# Proportion mature forest + openings edge density
mature.plus.openings.edge.model <- function(df) {
  glm(cbind(years.detect, years.no.detect) ~ proportion.mature + gap.edge.density, data=df, family=binomial(logit))
}

# Proportion mature forest + mature forest edge density.
mature.plus.mature.edge.model <- function(df) {
  glm(cbind(years.detect, years.no.detect) ~ proportion.mature + mature.edge.density, data=df, family=binomial(logit))
}

# Proportion high canopy cover.
proportion.high.canopy.model <- function(df) {
  glm(cbind(years.detect, years.no.detect) ~ proportion.canopy.high, data=df, family=binomial(logit))
}

# Proportion high canopy cover + high canopy fragmentation.
high.canopy.plus.canopy.edge.model <- function(df) {
  glm(cbind(years.detect, years.no.detect) ~ proportion.canopy.high + canopy.high.edge.density, data=df, family=binomial(logit))
}

# Null.
null.model <- function(df) {
  glm(cbind(years.detect, years.no.detect) ~ 1, data=df, family=binomial(logit))
}

# Nest the data frame.
nf <- occupancy.data %>% group_by(size) %>% nest()

# Apply the models.
nf <- nf %>% 
  mutate(
    # Suitability models.
    m.proportion.suitable=map(data, proportion.suitable.model),
    m.suitable.plus.suitable.edge=map(data, suitable.plus.suitable.edge.model),
    m.suitable.plus.hsi.edge=map(data, suitable.plus.hsi.edge.model),
    # Mature forest models.
    m.proportion.mature=map(data, proportion.mature.model),
    m.mature.plus.openings.edge=map(data, mature.plus.openings.edge.model),
    m.mature.plus.mature.edge=map(data, mature.plus.mature.edge.model),
    # Canopy cover models.
    m.proportion.high.canopy=map(data, proportion.high.canopy.model),
    m.high.canopy.plus.canopy.edge=map(data, high.canopy.plus.canopy.edge.model),
    # Null model.
    m.null=map(data, null.model)
         )
```

# Results

```{r aic-table, warning=FALSE, message=FALSE}
# Make an AIC table.
aictable <- nf %>% pivot_longer(-c(size, data), names_to='modname', values_to='model') %>% 
  group_by(size) %>% group_modify(~aictab(.x$model, modnames=.x$modname)) %>% 
  mutate(size=factor(size, levels=c('PFA', 'breeding area', 'home range', 'maximum range'))) %>% 
  arrange(size) %>% #filter(Delta_AICc <= 2) %>% 
  select(size, model=Modnames, AICc, K, delta.AICc=Delta_AICc, log.likelihood=LL)

# Count the number of models (rows) in each group.
n.pfa.models <- aictable %>% filter(size == 'PFA') %>% nrow()
n.ba.models <- aictable %>% filter(size == 'breeding area') %>% nrow()
n.hr.models <- aictable %>% filter(size == 'home range') %>% nrow()
n.mr.models <- aictable %>% filter(size == 'maximum range') %>% nrow()

aictable %>% kable() %>% 
  kable_styling() %>% 
  pack_rows(index=c('PFA'=n.pfa.models, 'Breeding area'=n.ba.models, 
                    'Home range'=n.hr.models, 'Maximum range'=n.mr.models))
```




