---
title: "R Notebook"
output: 
  bookdown::word_document2:
    fig_caption: true
---

```{r echo=FALSE, warning=FALSE, message=FALSE}
# Import conflict settings.
source('../src/conflicted.R')

# Load some libraries.
library(tidyverse)
library(lubridate)
library(sf)
library(adehabitatHR)
library(recurse)
library(flextable)
library(rmapshaper)
library(ggspatial)

# Copy out results from Bloxton.
blx <- tribble(
  ~class, ~family, ~category, ~common, ~prop.count,
  'Aves', 'Phasianidae', 'large bird', 'grouse', 31,
  'Aves', 'Columbidae', 'large bird', 'band-tailed pigeon', 22,
  'Aves', 'Corvidae', 'medium bird', "steller\'s jay", 14,
  'Aves', 'Picidae', 'medium bird', 'woodpecker', 4,
  'Aves', 'Turdidae', 'small bird', 'thrush', 4,
  'Aves', 'Unknown', 'small bird', 'unknown', 2,
  'Mammalia', 'Leporidae', 'large mammal', 'snowshoe hare', 10,
  'Mammalia', 'Sciuridae', 'medium mammal', 'douglas squirrel', 6,
  'Mammalia', 'Sciuridae', 'small mammal', 'flying squirrel', 4,
  'Mammalia', 'Rodentia', 'small mammal', 'unknown', 3
)

# This assumes "other birds" and "other mammals" are "small" sized, and splits woodpeckers and thrushes (which he gives as a combine 8%) into 4% each.

# Copy out results from Thraikill.
thr <- tribble(
  ~class, ~family, ~category, ~common, ~prop.count,
  'Aves', 'Phasianidae', 'large bird', 'ruffed grouse', 45,
  'Aves', 'Corvidae', 'medium bird', "steller\'s jay", 13,
  'Aves', 'Turdidae', 'medium bird', 'american robin', 13,
  'Aves', 'Phasianidae', 'large bird', 'ring-necked pheasant', 8,
  'Aves', 'Odontophoridae', 'large bird', 'mountain quail', 5,
  'Mammalia', 'Sciuridae', 'medium mammal', 'douglas squirrel', 13,
  'Mammalia', 'Aplodontiidae', 'large mammal', 'mountain beaver', 3
)

# Copy out results from Ethier.
eth <- tribble(
  ~class, ~family, ~category, ~common, ~frequency,
  'Aves', 'Turdidae', 'medium bird', 'varied thrush', 39,
  'Aves', 'Corvidae', 'medium bird', "steller\'s jay", 38,
  'Aves', 'Picidae', 'medium bird', 'northern flicker', 34,
  'Aves', 'Alcidae', 'large bird', 'marbled murrelet', 15,
  'Aves', 'Columbidae', 'large bird', 'band-tailed pigeon', 5,
  'Aves', 'Corvidae', 'medium bird', 'gray jay', 4,
  'Aves', 'Hirundinidae', 'small bird', 'barn swallow', 2,
  'Aves', 'Strigidae', 'medium bird', 'northern pygmy owl', 1,
  'Aves', 'Turdidae', 'small bird', 'swainson\'s thrush', 1,
  'Aves', 'Tyrannidae', 'small bird', 'western wood-peewee', 1,
  'Mammalia', 'Sciuridae', 'medium mammal', 'red squirrel', 69,
  'Mammalia', 'Cricetidae', 'small mammal', 'vole', 1,
  'Mammalia', 'Unknown', 'small mammal', 'unknown', 1
)

# Ethier's results are WRONG!!!!! Must convert to percents.
eth <- eth %>% 
  mutate(total.count=sum(frequency), prop.count=frequency/total.count*100) %>% 
  select(-c(frequency, total.count))

# Add author information to each data frame.
blx <- blx %>% mutate(author='bloxton')
thr <- thr %>% mutate(author='thraikill')
#eth <- eth %>% mutate(author='ethier')

# Join the two frames together.
rv <- bind_rows(blx, thr)

# Add mass.
# rv <- left_join(rv, prey.list) %>% mutate_if(is.character, replace_na, 'Unknown') %>% 
#   left_join(mean.mass) %>% 
#   mutate(mass=coalesce(mass, average)) %>% 
#   select(author, class, family, order, genus, species, binomial, common, category, prop.count, mass)

# Calculate biomass.
#rv <- rv %>% mutate(biomass=prop.count*mass)

# Calculate counts from my own data.
# prop.count.avian.pr <- filter(diet.items, source != 'C') %>% 
#   mutate(total=n()) %>% filter(class == 'Aves') %>% 
#   mutate(count=n(), to.count=count/total*100) %>% 
#   distinct(to.count) %>% as.numeric()
# 
# prop.count.mammal.pr <- filter(diet.items, source != 'C') %>% 
#   mutate(total=n()) %>% filter(class == 'Mammalia') %>% 
#   mutate(count=n(), to.count=count/total*100) %>% 
#   distinct(to.count) %>% as.numeric()

# Add in new data.
mn <- tribble(
  ~author, ~class, ~proportion,
  'case', 'Aves', 60.55,
  'case', 'Mammalia', 39.45,
  'watson', 'Aves', 53,
  'watson', 'Mammalia', 47,
  'lewis', 'Aves', 83.65,
  'lewis', 'Mammalia', 16.35
)

# This is combined pellets-and-remains, for my own data as well as Watson's.

# Join with Bloxton & Thraikill.
pnw.class.summary <- rv %>% group_by(author, class) %>% 
  mutate(proportion=sum(prop.count)) %>% 
  distinct(author, class, proportion) %>% 
  bind_rows(mn) %>% 
  pivot_wider(names_from=class, values_from=proportion) %>% 
  arrange(Aves)

# For squirrels, also.
# prop.count.squirrel.pr <- filter(diet.items, source != 'C') %>% 
#   mutate(total=n()) %>% filter(group == 'squirrel') %>% 
#   mutate(count=n(), to.count=count/total*100) %>% 
#   distinct(to.count) %>% as.numeric()

mn.sq <- tribble(
  ~author, ~genus, ~prop.count,
  'case', 'Tamiasciurus', 24.22,
  'watson', 'Tamiasciurus', 15.1,
)

pnw.squirrel.summary <- rv %>% filter(common %in% c('douglas squirrel', 'red squirrel')) %>% 
  select(genus=common, prop.count, author) %>%
  bind_rows(mn.sq) %>% 
  arrange(prop.count)
```

Tree squirrels are the same key prey identified on Vancouver Island (Ethier 1999) but not from other regions of the Pacific Northwest, where the key prey is generally grouse (Watson et al. 1998, Thraikill et al. 2000, Bloxton 2002, Lewis et al. 2006). The key prey from my study area is more similar to those identified elsewhere in western North America, where the most important prey species is also mammalian, generally either a sciurid or a leporid. This unexpected result may be in part the product of methodological differences between my study and others conducted in the Pacific Northwest. When the results from studies across temperate rainforest ecosystems are standardized (data from pooled pellets-and-remains or remains only, measured by counts), the difference between British Columbia and other regions in the Pacific Northwest is much less pronounced. However, the proportion of mammalian prey in the diet, particularly tree squirrels, remains markedly higher within my study area see (Figure \@ref(fig:density-map)). This is likely the result of relatively high tree squirrel abundance within my study area, which is much greater (Ransome & Sullivan 2003) than in other temperate rainforest ecosystems (reviewed in Carey et al. 1995). No Pacific Northwest study has assessed goshawk diet and and absolute prey abundance simultaneously (but see Ethier 1999), but regional data on tree squirrel abundance hint that, as abundance varies, so may goshawk specialization. Across two ecological zones present in my study area I observed only minor variation in goshawk diet and no variation in the dominance of tree squirrels in the diet, indicating a slight difference in the prey community of these two zones but a similar abundance of tree squirrels. Overall, these results suggest goshawks in my study area pursue a specialized generalist (Elmhagen et al. 2000) foraging strategy--a generalist predator opportunistically exploiting a locally abundant prey source.

```{r echo=FALSE}
# Copy out approximate coordinate for all the study areas.
squirrel.coords <- tribble(
  ~lon, ~lat, ~author,
  #-125.621, 50.023, 'ethier',
  -122.019, 46.326, 'watson',
  -122.469, 49.588, 'case',
  -124.128, 47.999, 'carey95',
  -123.755, 47.464, 'bloxton',
  -123.842, 44.483, 'thraikill',
  -123.798, 42.685, 'carey92',
  -122.965, 49.335, 'ransome',
  -121.463, 48.614, 'carey93'
)

# Copy out squirrel densities (animals/ha)
squirrel.density <- tribble(
  ~type, ~value, ~author,
  'density', 0.2, 'carey92',
  'density', 0.1, 'carey95',
  'density', 0.5, 'carey93',
  'density', 1.1, 'ransome'
)

# Make the data set.
squirrel.map <- pnw.squirrel.summary %>% 
  mutate(genus='diet') %>% 
  rename(type=genus, value=prop.count) %>% 
  bind_rows(squirrel.density) %>% 
  left_join(squirrel.coords, by=c('author'))

# Make a box for map size.
squirrel.box <- squirrel.coords %>% 
  st_as_sf(coords=c('lon', 'lat')) %>% 
  st_bbox()

# Bring in some base data.
n.america <- read_sf(dsn='../data/external/ne_10m_land.shp', layer='ne_10m_land')
rivers <- read_sf(dsn='../data/external/ne_10m_rivers_lake_centerlines.shp', 
                  layer='ne_10m_rivers_lake_centerlines')
```

```{r density-map, echo=FALSE, fig.cap="this is a caption"}
# Make a map.
ggplot() +
  geom_sf(data=n.america, fill='lightgrey') +
  #geom_sf(data=sc.region, aes(fill='darkgrey')) +
  geom_sf(data=rivers) +
  coord_sf(xlim=c(st_bbox(squirrel.box)[1] - 0.75, st_bbox(squirrel.box)[3] + 0.75), 
           ylim=c(st_bbox(squirrel.box)[2] - 0.25, st_bbox(squirrel.box)[4] + 0.25)) +
  geom_point(data=squirrel.map, aes(x=lon, y=lat, shape=type)) +
  geom_label(data=squirrel.map, 
             aes(x=lon, y=lat, label=value, color=type), nudge_x=0.6, nudge_y=-0.1) +
  theme_void()
```
