---
title: "VRI models"
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, message=FALSE, warning=FALSE)
```

Same thing as before, but with VRI instead of BEC data. This is capturing more fine-scale forest attributes.

```{r}
# Import conflict settings.
source('../src/conflicted.R')

# Load some libraries.
library(tidyverse)
library(raster)
library(sf)
library(landscapemetrics)
library(vegan)
library(AICcmodavg)

# Import VRI shapefile.
shp.vri <- st_read('../data/external/VRI_sc_all/VRI_lower-mainland.shp')
```
Now to do some land cover classification. Last time I did something pretty simple, but I think I'd like a little more nuanced this time.

* Undefined
* Bare ground/rock
* Ocean
* Freshwater
* Alpine
* Wetland
* Shrub & burns (ie natural openings)
* Deciduous forest
* Regen/clearcut forest (<20)
* Young forest (20-70)
* Mature forest (70-250)
* Old forest (>250)

Louise does >250, 140-250, and <140. But that makes sense for MAMU.
The HSI tops out at 150.
McClaren says old > 250, mature/second growth 40-140

```{r}
# Define class levels.
vri.landcover.levels <- data.frame(ID=0:11, class.name=
                                     c('undefined', 'rock', 'ocean', 'freshwater',
                                       'alpine', 'wetland',
                                       'shrub', 'deciduous', 'regen',
                                       'young', 'mature', 'old'))
# Add to VRI.
shp.vri <- shp.vri %>% mutate(land.cover=case_when(
  # Bare ground/rock
  BCLCS_LV_1 == 'N' & BCLCS_LV_2 != 'W' ~ 1,
  # Ocean
  BCLCS_LV_2 == 'W' & BCLCS_LV_5 == 'OC' ~ 2 ,
  # Freshwater (rivers, lakes, & ponds)
  BCLCS_LV_2 == 'W' & BCLCS_LV_5 != 'OC' ~ 3,
  # Shrub
  BCLCS_LV_1 == 'V' & BCLCS_LV_2 == 'N' ~ 6,
  # Burns 
  BCLCS_LV_1 == 'N' & BCLCS_LV_5 == 'BU' ~ 6,
  # Mixed or deciduous trees
  BCLCS_LV_1 == 'V' & BCLCS_LV_4 %in% c('TB', 'TM') ~ 7,
  # Coniferous trees
  BCLCS_LV_4 == 'TC' & PROJ_AGE_1 < 20 ~ 8,
  BCLCS_LV_4 == 'TC' & PROJ_AGE_1 >= 20 & PROJ_AGE_1 < 70 ~ 9,
  BCLCS_LV_4 == 'TC' & PROJ_AGE_1 >= 70 & PROJ_AGE_1 < 250 ~ 10,
  BCLCS_LV_4 == 'TC' & PROJ_AGE_1 >= 250 ~ 11,
  # Alpine
  BCLCS_LV_3 == 'A' ~ 4,
  # Wetland
  BCLCS_LV_3 == 'W'~ 5,
  TRUE ~ 0
))
```


```{r eval=FALSE}
# Set raster extent based on VRI shapefile.
ext <- extent(shp.vri)

# Make an empty raster to populate with values.
r <- raster(ext, res=c(15, 15))

# Populate VRI polygon data onto empty raster grid.
r.vri <- rasterize(shp.vri, r, 'land.cover')

# Define land cover levels.
landcover.levels <- data.frame(ID=0:11, class.name=
                                     c('undefined', 'rock', 'ocean', 'freshwater',
                                       'alpine', 'wetland',
                                       'shrub', 'deciduous', 'regen',
                                       'young', 'mature', 'old'))

# Add levels to raster.
levels(r.vri) <- landcover.levels

# Assign CRS.
crs(r.vri) <- CRS('+proj=utm +zone=10 +datum=NAD83 +units=m +no_defs')

# And save.
writeRaster(r.vri, '../data/processed/vri_lower-mainland.tif', format='GTiff')
```

So that's land cover. Is there anything else I want to get from the VRI?

One idea is tree diversity, or dominant tree type. No point whatsoever in age or height. Also canopy closure, if there's enough coverage for it.

```{r}
# How good is my canopy closure data?

shp.vri %>% data.frame() %>% filter(BCLCS_LV_2 == 'T') %>% 
  mutate(forest.patches=n()) %>% 
  filter(is.na(CR_CLOSURE)) %>% 
  mutate(missing.closure=n()) %>% 
  select(forest.patches, missing.closure) %>% 
  distinct()
```
Ok, that's great, if only 4 patches labeled as forest are missing canopy closure information.