---
title: "R Notebook"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, message=FALSE, warning=FALSE)
```

I've run into a small problem with my BEC raster--some classes don't seem to be showing up properly in the analysis.

```{r}
# Load up some libraries.
library(tidyverse)
library(sf)
library(raster)
library(landscapemetrics)

# Import nests and calculate centroids.
cam.sites <- read_csv('../data/processed/the_big_list_of_nests.csv') %>% 
  group_by(name) %>% 
  mutate_at(c('lat', 'lon'), mean) %>% 
  mutate_at(vars(starts_with('status')), max) %>% 
  mutate_at(c('telemetry', 'cameras', 'remains'), max) %>% 
  dplyr::select(-nest, -NOTES) %>% 
  distinct() %>% 
  filter(cameras > 0)

# Drop TCR, ungroup.
cam.sites <- cam.sites %>% filter(site != 'TCR') %>% 
  ungroup() %>% 
  rownames_to_column(var='plot_id') %>% 
  mutate(plot_id=as.integer(plot_id))

# Make site table a spatial object and make it UTMs.
cam.sites.sf <- ungroup(cam.sites) %>% st_as_sf(coords=c('lon', 'lat')) %>%
  st_set_crs('+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs') %>%
  st_transform("+proj=utm +zone=10 +datum=WGS84 +units=m +no_defs")

# Convert approximate homerange area in ha to radius in m
a.hr.ha <- 3700
r.hr.m <- sqrt(a.hr.ha*10000/pi)

# Import the BEC raster.
r.bec <- raster('../data/interim/bec_raster_SC.tif')
```

That's just the basics. So far, so good. What are the levels on the BEC raster, as it stands?

```{r}
levels(r.bec)
```

Ah, so it already has levels assigned. I tried to re-assign them. What did I try to re-assign them to?

```{r}
# Import the original shapefile.
shp.bec <- st_read('../data/processed/BEC_annual-homerange_clipped.shp')

# Get BEC levels.
bec.levels <- data.frame((shp.bec[["MAP_LABEL"]])) %>% 
  rownames_to_column() %>% 
  rename(ID=1, map_label=2) %>% 
  mutate(ID=as.numeric(ID))

bec.levels
```

Verrrryyy different. The question is, which ones are right?


```{r}
# Build sample.
my.plots <- data.frame(construct_buffer(coords=cam.sites.sf, shape='circle',
                                            size=r.hr.m, return_sp=FALSE))

# Make the graph.
bec.plot <- raster::as.data.frame(r.bec, xy=TRUE) %>% 
  ggplot() +
  geom_raster(aes(x=x, y=y, fill=factor(bec_raster_SC_category)))

# Add sample polygons.
bec.plot +
  geom_polygon(data=my.plots, aes(x = my.plots[, 1],
                                  y = my.plots[, 2],
                                  group = my.plots[, 3]),
                                  col = "black", fill = NA)
```

That's with the built-in levels. What about the levels I added?

```{r}
bad.bec <- r.bec

levels(bad.bec) <- bec.levels

bad.plot <- raster::as.data.frame(bad.bec, xy=TRUE) %>% 
  ggplot() +
  geom_raster(aes(x=x, y=y, fill=factor(bec_raster_SC_map_label)))

# Add sample polygons.
bad.plot +
  geom_polygon(data=my.plots, aes(x = my.plots[, 1],
                                  y = my.plots[, 2],
                                  group = my.plots[, 3]),
                                  col = "black", fill = NA)
```

So it looks like adding levels was a mistake. But can I get a closer look?

```{r}
# Pull out a toy known to be a problem.
utz.coord <- cam.sites.sf %>% filter(site == 'UTZ')

utz.plot <- data.frame(construct_buffer(coords=utz.coord, shape='circle',
                                            size=r.hr.m, return_sp=FALSE))

# Get bounding coordinates for site.
utzxmin <- st_bbox(utz.coord)[1] - r.hr.m
utzymin <- st_bbox(utz.coord)[2] - r.hr.m
utzxmax <- st_bbox(utz.coord)[3] + r.hr.m
utzymax <- st_bbox(utz.coord)[4] + r.hr.m

# Add to graph.
bec.plot +
  geom_polygon(data=utz.plot, aes(x = utz.plot[, 1],
                                  y = utz.plot[, 2],
                                  group = utz.plot[, 3]),
                                  col = "black", fill = NA) +
  xlim(utzxmin, utzxmax) +
  ylim(utzymin, utzymax)
```

And with the added labels...

```{r}
# Using bad bec.
bad.plot +
  geom_polygon(data=utz.plot, aes(x = utz.plot[, 1],
                                  y = utz.plot[, 2],
                                  group = utz.plot[, 3]),
                                  col = "black", fill = NA) +
  xlim(utzxmin, utzxmax) +
  ylim(utzymin, utzymax)
```

Ohhhhhh super different. So that really is where the problem is.

As long as I'm here, can I figure out a way to plot all of my sites together?
```{r}
# Site 1
bec.plot +
  geom_polygon(data=my.plots, aes(x = my.plots[, 1],
                                  y = my.plots[, 2],
                                  group = my.plots[, 3]),
                                  col = "black", fill = NA) +
  xlim(st_bbox(cam.sites.sf[1, ])[1] - r.hr.m, st_bbox(cam.sites.sf[1, ])[3] + r.hr.m) +
  ylim(st_bbox(cam.sites.sf[1, ])[2] - r.hr.m, st_bbox(cam.sites.sf[1, ])[4] + r.hr.m) +
  theme_classic() + labs(title='MTF')

# Site 2
bec.plot +
  geom_polygon(data=my.plots, aes(x = my.plots[, 1],
                                  y = my.plots[, 2],
                                  group = my.plots[, 3]),
                                  col = "black", fill = NA) +
  xlim(st_bbox(cam.sites.sf[2, ])[1] - r.hr.m, st_bbox(cam.sites.sf[2, ])[3] + r.hr.m) +
  ylim(st_bbox(cam.sites.sf[2, ])[2] - r.hr.m, st_bbox(cam.sites.sf[2, ])[4] + r.hr.m) +
  theme_classic() + labs(title='MTC')

# Site 3
bec.plot +
  geom_polygon(data=my.plots, aes(x = my.plots[, 1],
                                  y = my.plots[, 2],
                                  group = my.plots[, 3]),
                                  col = "black", fill = NA) +
  xlim(st_bbox(cam.sites.sf[3, ])[1] - r.hr.m, st_bbox(cam.sites.sf[3, ])[3] + r.hr.m) +
  ylim(st_bbox(cam.sites.sf[3, ])[2] - r.hr.m, st_bbox(cam.sites.sf[3, ])[4] + r.hr.m) +
  theme_classic() + labs(title='RLK')

# Site 4
bec.plot +
  geom_polygon(data=my.plots, aes(x = my.plots[, 1],
                                  y = my.plots[, 2],
                                  group = my.plots[, 3]),
                                  col = "black", fill = NA) +
  xlim(st_bbox(cam.sites.sf[4, ])[1] - r.hr.m, st_bbox(cam.sites.sf[4, ])[3] + r.hr.m) +
  ylim(st_bbox(cam.sites.sf[4, ])[2] - r.hr.m, st_bbox(cam.sites.sf[4, ])[4] + r.hr.m) +
  theme_classic() + labs(title='TMC')

# Site 5
bec.plot +
  geom_polygon(data=my.plots, aes(x = my.plots[, 1],
                                  y = my.plots[, 2],
                                  group = my.plots[, 3]),
                                  col = "black", fill = NA) +
  xlim(st_bbox(cam.sites.sf[5, ])[1] - r.hr.m, st_bbox(cam.sites.sf[5, ])[3] + r.hr.m) +
  ylim(st_bbox(cam.sites.sf[5, ])[2] - r.hr.m, st_bbox(cam.sites.sf[5, ])[4] + r.hr.m) +
  theme_classic() + labs(title='UTZ')
```
