---
title: "Univariables"
output: html_notebook
---

```{r options, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE)
```

See what I did there? Let's look at all the landscape variables used in the landscape analysis.

```{r message=FALSE}
# Import conflict settings.
source('../src/conflicted.R')

#Load some libraries.
library(tidyverse)
library(vegan)
library(raster)
library(sf)
library(landscapemetrics)
library(broom)
library(knitr)
library(kableExtra)
library(extrafont)
```

Since this is just for making pretty pictures, I'll take a shortcut and bring in some previously calculated data.

```{r}
# Load the data.
data <- readRDS('../data/interim/occupancy_landscape_metrics.rds')

# Filter out the problematic Stewart Island site.
data <- data %>% filter(site != 'SIS')

# Load all the rasters.
source('../src/load_rasters.R')

# Pull out one site as a demo.
utz <- centroids %>% filter(site == 'UTZ') %>% select(MEAN_X, MEAN_Y)
```

Then start making a lot of graphs.

## Forest type

```{r}
# Convert raster to data frame for easier plotting.
df.bec <- raster::as.data.frame(r.bec, xy=TRUE)

# Plot a demo.
ggplot(df.bec, aes(x=x, y=y, fill=bec_raster_full_category)) +
  geom_raster() +
  xlim(utz$MEAN_X-7050, utz$MEAN_X+7050) +
  ylim(utz$MEAN_Y-7050, utz$MEAN_Y+7050) +
  labs(title='Forest types at site UTZ') +
  theme_void()

```

### Diversity of forest types

The diversity of BEC subzones/variants within the landscape.

```{r warning=FALSE}
# Show for all sites.
select(bec.landscape.metrics, nest, radius, bec.diversity) %>% 
  ggplot(aes(x=factor(radius), y=bec.diversity, fill=factor(radius))) +
  geom_boxplot() +
  geom_jitter() +
  scale_fill_brewer(palette='Set2') +
  theme_classic() +
  scale_x_discrete(labels=c('PFA', 'Breeding area', 'Breeding home range', 'maximum range')) +
  theme(axis.title.x=element_blank(), legend.position='none') +
  labs(title='Diversity of forest types') +
  theme(text=element_text(family="Lato Semibold"))
```

## Land cover

```{r}
# Convert raster to data frame for easier plotting.
df.landcover <- raster::as.data.frame(r.landcover, xy=TRUE)

# Plot a demo.
ggplot(df.landcover, aes(x=x, y=y, fill=bec_raster_full_category)) +
  geom_raster() +
  xlim(utz$MEAN_X-7050, utz$MEAN_X+7050) +
  ylim(utz$MEAN_Y-7050, utz$MEAN_Y+7050) +
  labs(title='Forest types at site UTZ') +
  theme_void()

```

### Diversity of forest types

The diversity of BEC subzones/variants within the landscape.

```{r warning=FALSE}
# Show for all sites.
select(bec.landscape.metrics, nest, radius, bec.diversity) %>% 
  ggplot(aes(x=factor(radius), y=bec.diversity, fill=factor(radius))) +
  geom_boxplot() +
  geom_jitter() +
  scale_fill_brewer(palette='Set2') +
  theme_classic() +
  scale_x_discrete(labels=c('PFA', 'Breeding area', 'Breeding home range', 'maximum range')) +
  theme(axis.title.x=element_blank(), legend.position='none') +
  labs(title='Diversity of forest types') +
  theme(text=element_text(family="Lato Semibold"))
```













