---
title: "Camera data collection methods"
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE)
```

```{r}
# Load up some libraries.
library('tidyverse')
library('lubridate')
library('ggplot2')
library('knitr')
library('kableExtra')

# Import the data.
df <- read.csv('../data/interim/camera_corrected.csv', stringsAsFactors=FALSE)

# Do the datetime thing.
df <- df %>% mutate(datetime=parse_date_time(datetime, 
                       orders=c('%Y-%m-%d %H:%M:%S', '%Y/%m/%d %H:%M:%S')))

n.nests <- df %>% distinct(site) %>% summarize(n()) %>% as.numeric()

install.date <- df %>% group_by(site) %>% summarize(d=min(datetime)) %>% 
  mutate(day=day(d), m=month(d))
```

**Nest Cameras**

We quantified the diet of breeding goshawks using digital trail cameras placed at `r n.nests` nests during 2019. Nests were selected for camera installation based on timing of discovery, ease of access, and the presence of suitable trees. Each camera was attached to the trunk of a tree adjacent to the nest tree, approximately five meters away from and slightly above the nest itself. Cameras (Reconyx, UltraFire and HyperFire models) were programmed to take three photos one second apart when triggered by motion, and an additional one photo every thirty minutes. Installation took place during the early nestling phase
(between `r install.date %>% filter(d==min(d)) %>% dplyr::select(day) %>% as.numeric()` 
`r install.date %>% filter(d==min(d)) %>% dplyr::select(m) %>% as.numeric() %>% month(label=TRUE, abbr=FALSE)`
and `r install.date %>% filter(d==max(d)) %>% dplyr::select(day) %>% as.numeric()` `r install.date %>% filter(d==min(d)) %>% dplyr::select(m) %>% as.numeric() %>% month(label=TRUE, abbr=FALSE)`) and cameras were left in place until after juvenile dispersal.

Chicks were aged from photos taken shortly after camera installation using a pictoral guide (Boal 1994) and each nest was assigned a hatch date based on median chick age. Nest productivity was defined as the number of chicks per nest that survived to 35 days of age (Woodbridge and Hargis 2006). <!-- Check that citation! -->

**Prey Identification**

We identified prey items delivered to the nest following Lewis et al. (2004). Nest camera photos were reviewed and each new prey item was recorded and identified to species when possible, using a combination of study skins, field guides, and researcher knowledge. When identification to species was not possible, items were identified to the lowest possible taxonomic level. Items were additionally categorized by size (small, medium, or large) following Miller et al. (2014). As very few items could be successfully aged, we did not include prey age in our analysis.

```{r}
# Import prey items data
source('../src/prey_attributes.R')
```

Prey items identified to species were assigned mass using Nagorsen (2002; mammals) and Billerman et al. (2020; birds). When unable to differentiate between species within a single genus (such as *Eutamias* and *Myotis*), we assigned mass by averaging the masses of all possible species, determined by range maps. Unidentified items and partial items were assigned mass by averaging the masses of the identified species in that size and taxonomic group. For example, an unidentified sciurid would receive the average mass of all identified sciurid species, while a small portion of an unidentified bird would receive the average mass of all identified small bird species.

We calculated the relative proportions of avian and mammalian biomass delivered to all nests during the study period. For each nest, we calculated the mean prey deliveries per day by count and by biomass. <!-- Is it clear that this is the average of an average? --> Daily biomass for all six nests was pooled to determine the effect of brood size and brood age on delivery rate. We transformed biomass data by taking the natural log of daily biomass delivered to each nest and used linear models to assess the effects of brood age and number of chicks present in the nest on deliveries.

We calculated prey species diversity for the entire study area and for each nest using items identified to genus or species using Simpson's Diversity Index (citation needed). We calculated dietary overlap between nests using Morisita's Index of Similarity (Krebs 1998). <!-- but check that citation -->

We report all means with $\pm$ SD.

## Results

```{r}
# Import basic camera nest data
cameras <- read.csv('../data/raw/productivity_2019.csv', stringsAsFactors=FALSE)

# Do the date conversion thing.
cameras <- cameras %>% mutate(date=ymd(date), branch=ymd(branch), fledge=ymd(fledge))

# How many fledged?
n.fledge <- cameras %>% group_by(n_fledge) %>% 
  summarize(n=n())

# How many deliveries?
n.del <- df %>% filter(interest %in% c('newprey', 'delivery')) %>% nrow()

# How many identified to class?
n.to.class <- items %>% filter(class != 'Unknown') %>% nrow()

# How many idetnified to genus?
n.to.genus <- items %>% filter(genus != 'Unknown') %>% nrow()
```

We observed no nest abandonment following camera installation. One nest failed 9 days following camera install, while the other five nests successfully fledged at least one chick. Successfully nests fledged 1 (*n* = `r as.numeric(n.fledge[2, 2])`), 2 (*n* = `r as.numeric(n.fledge[3, 2])`), or 3 (*n* = `r as.numeric(n.fledge[4, 2])`) chicks. The unsuccessful nest failed after two chicks succumbed to siblicide and the third appeared to fledge prematurely, though the exact cause of failure is unknown. Two other nests were observed to lose a single chick each due to apparent siblicide. <!-- Include site names? --> <!-- Include chick ages -->

<!-- Fledge dates? -->

**Prey deliveries**

We obtained `r nrow(df)` photos from `r nrow(cameras)` nests during the 2019 breeding season. A total of `r n.del` prey item deliveries were recorded. `r round((n.del-nrow(items))/n.del*100, digits=0)`% of items were obscured from the camera during delivery and consumption and were removed from the analysis. Out of the `r nrow(items)` visible items, `r round(n.to.class/nrow(items)*100, digits=0)`% were identified to class and `r round(n.to.genus/nrow(items)*100, digits=0)`% to genus or species. Small and medium birds were disproportionately represented among unidentified items, frequently arriving at the nest already plucked and decapitated.

Across the entire study area, we observed `r items %>% filter(genus != 'Unknown') %>% dplyr::select(genus, species) %>% distinct() %>% summarize(n()) %>% as.numeric()` different prey species delivered to nests. By biomass, mammals made up the largest proportion of deliveries (`r items %>% mutate(mass=as.numeric(mass), t.mass=sum(mass)) %>% filter(class == 'Mammalia') %>% mutate(m.mass=sum(mass), p.mass.m=(m.mass/t.mass)*100) %>% dplyr::select(p.mass.m) %>% distinct() %>% as.numeric() %>% round()`%). This was due to the overwhelming number of tree squirrels (*Tamiasciurus* spp.) delivered to nests, which provided `r items %>% mutate(mass=as.numeric(mass), t.mass=sum(mass)) %>% filter(genus == 'Tamiasciurus') %>% mutate(sq.mass=sum(mass), p.mass.sq=(sq.mass/t.mass)*100) %>% dplyr::select(p.mass.sq) %>% distinct() %>% as.numeric() %>% round()`% of biomass. Birds made up `r items %>% mutate(mass=as.numeric(mass), t.mass=sum(mass)) %>% filter(class == 'Aves') %>% mutate(a.mass=sum(mass), p.mass.a=(a.mass/t.mass)*100) %>% dplyr::select(p.mass.a) %>% distinct() %>% as.numeric() %>% round()`% of the diet, with the final `r items %>% mutate(mass=as.numeric(mass), t.mass=sum(mass)) %>% filter(class == 'Unknown') %>% mutate(u.mass=sum(mass), p.mass.u=(u.mass/t.mass)*100) %>% dplyr::select(p.mass.u) %>% distinct() %>% as.numeric() %>% round()`% of prey biomass unable to be identifed as either bird or mammal.

```{r}
# Number (count) of deliveries per day.
del.per.day <- df %>% filter(interest %in% c('delivery', 'newprey')) %>% 
  mutate(day=date(datetime)) %>% 
  group_by(site) %>% 
  mutate(first=min(day), last=max(day), n.day=last-first) %>% 
  add_count(site, day, name='n.del') %>% 
  add_count(site, name='t.del') %>% 
  mutate(n.day=as.numeric(n.day), per.day=t.del/n.day) %>% 
  dplyr::select(site, day, n.day, n.del, t.del) %>% 
  distinct() %>% 
  complete(nesting(site), day=seq(min(day), max(day), by='day')) %>% 
  fill(n.day, t.del) %>% 
  mutate(n.del=replace_na(n.del, 0))

source('../src/prey_attributes_date.R')

# Amount (biomass) of deliveries per day.
mass.by.day <- items.alt %>% mutate(date=date(datetime)) %>% 
  group_by(site, date) %>% 
  mutate(d.mass=sum(mass)) %>% 
  dplyr::select(site, date, d.mass) %>% 
  arrange(site, date) %>% 
  distinct() %>% 
  ungroup() %>% group_by(site) %>% 
  complete(nesting(site), date=seq(min(date), max(date), by='day')) %>% 
  mutate(d.mass=replace_na(d.mass, 0))

# Import nest data.
nests <- read.csv('../data/raw/camera_nests_2019.csv', stringsAsFactors=FALSE)

# Do the datetime thing.
nests <- nests %>% mutate(est_hatch=ymd(est_hatch))

# Table with date and age for each site.
age <- nests %>% dplyr::select(site, est_hatch) %>% 
  mutate(fledge=est_hatch + 99) %>% 
  pivot_longer(cols=2:3) %>% 
  group_by(site) %>% 
  complete(nesting(site), value=seq(min(value), max(value), by='day')) %>% 
  mutate(age=seq(1:100), date=value) %>% 
  dplyr::select(site, date, age)

# Re-digest mass data into model-friendly format.
mass.for.model <- items.alt %>% mutate(date=date(datetime)) %>% 
  group_by(site, date) %>% 
  mutate(d.mass=sum(mass), log.d.mass=log(d.mass)) %>% 
  pivot_wider(names_from=class, values_from=mass, values_fill=list(mass=0)) %>% 
  mutate(Aves=sum(Aves), Mammalia=sum(Mammalia), Unknown=sum(Unknown)) %>% 
  arrange(site, date) %>% 
  dplyr::select(site, date, Aves, Mammalia, Unknown, d.mass, log.d.mass) %>% 
  distinct() %>% 
  mutate(p.mammal=Mammalia/d.mass)

# Now graft them together.
mass.for.model <- left_join(mass.for.model, age) %>% 
  mutate(age=as.numeric(age))

# And make a model.
m.mass.by.age <- lm(log.d.mass ~ age, data=mass.for.model)
```

Prey deliveries averaged `r del.per.day %>% group_by(site) %>% summarize(mean=mean(n.del)) %>% summarize(mean(mean)) %>% as.numeric() %>% round(digits=2)` $\pm$ `r del.per.day %>% group_by(site) %>% summarize(mean=mean(n.del)) %>% summarize(sd(mean)) %>% as.numeric() %>% round(digits=2)` deliveries/day, although all nests observed occasionally went at least one or more days without any deliveries. Average biomass of items was `r round(mean(items$mass), digits=2)` $\pm$ `r round(sd(items$mass), digits=2)` g, and the average daily biomass of prey delivered to each nest was `r mass.by.day %>% group_by(site) %>% summarize(mean=mean(d.mass)) %>% summarize(mean(mean)) %>% as.numeric() %>% round(digits=2)` $\pm$ `r mass.by.day %>% group_by(site) %>% summarize(mean=mean(d.mass)) %>% summarize(sd(mean)) %>% as.numeric() %>% round(digits=2)` g/day. Brood age did not affect delivery rates, with older broods receiving approximately the same biomass per day as younger broods (*P* = `r summary(m.mass.by.age)$coefficients[2,4] %>% round(digits=2)`). Brood size did not affect delivery rates, with large broods receiving the same daily biomass per chick as smaller broods.

The proportion of prey biomass made of mammalian prey increased with brood age until age x, and then slowly decreased. Brood size also had an affect on the proportion of mammalian prey, with larger broods receiving a greater proportion of mammalian biomass.

The overall index of diversity for the diet of all x nests was x. For individual nests, diversity ranged from x to x (mean=x). Overlap was consistently low, ranging from x to x (mean=x). There was a slight but non-significant difference in diversity between transition zone and coastal zone nests. Overlap was higher within zones than between zones.