---
title: "Diet, revisited"
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, message=FALSE, warning=FALSE)
```

1. Add missing models are re-run.
    * Prerequisites:
      * ~~choose canopy cover categories~~
      * ~~make canopy cover rasters~~
      * ~~make older forest rasters~~
2. Fix some stuff I've been doing wrong.
    * remake landcover rasters at 100m scale
    * rerun diet breakdown with counts
    Next steps:
    * compare counts and biomass breakdowns
3. Run the models also for physical specimen data.
    * Prerequisites:
      * finish cleaning physical data
      * determine if there's a cutoff for sample size
    * Next steps:
      * compare results from camera and physical data
4. Compare diet between coastal and transition zones.
    * Prerequisites:
      * assign each site to cs or tz
    * Next steps:
      * compare results from camera and physical data

```{r}
# Import conflict settings.
source('../src/conflicted.R')

# Load some libraries.
library(tidyverse)
library(lubridate)
library(vegan)
library(ggplot2)

# Bring in data.
source('../src/prey_attributes_revised.R')
```

Compare biomass and counts for cameras.

```{r}
# Calculate biomass and counts for mammalian prey.
proportion.mammal.cameras <- diet.items %>% 
  filter(method == 'camera') %>% 
  group_by(nest) %>% 
  mutate(total.mass=sum(mass), total.count=n()) %>% 
  filter(class == 'Mammalia') %>% 
  mutate(mass.mm=sum(mass), count.mm=n(), 
         prop.mm.mass=mass.mm/total.mass, prop.mm.count=count.mm/total.count) %>% 
  select(nest, prop.mm.mass, prop.mm.count) %>% distinct() %>% 
  pivot_longer(!nest, names_to='metric', values_to='value')

# Also in wide format.
proportion.mammal.cameras.wide <- proportion.mammal.cameras %>% 
  pivot_wider(names_from=metric, values_from=value)

proportion.mammal.cameras.wide

# Look at the distribution.
qqnorm(proportion.mammal.cameras$value)
qqline(proportion.mammal.cameras$value)

# Check for normality.
shapiro.test(proportion.mammal.cameras$value)
```

So.... sort of normal-ish? I'll give it a shot. Is there is a significant difference between using counts or biomass?

```{r}
# Run a t-test.
t.test(proportion.mammal.cameras.wide$prop.mm.mass, proportion.mammal.cameras.wide$prop.mm.count)
```

So the two methods (biomass vs. counts) do give slightly different results. Which is interesting, because Lewis says they don't. Generally, counts are more widely reported.

```{r}
# Do the same with proportion avian biomass/counts.

# Calculate biomass and counts for mammalian prey.
proportion.avian.cameras <- diet.items %>% 
  filter(method == 'camera') %>% 
  group_by(nest) %>% 
  mutate(total.mass=sum(mass), total.count=n()) %>% 
  filter(class == 'Aves') %>% 
  mutate(mass.av=sum(mass), count.av=n(), 
         prop.av.mass=mass.av/total.mass, prop.av.count=count.av/total.count) %>% 
  select(nest, prop.av.mass, prop.av.count) %>% distinct() 

proportion.avian.cameras

# Also in long format.
proportion.avian.cameras.long <- proportion.avian.cameras %>% 
  pivot_longer(!nest, names_to='metric', values_to='value')

# Look at the distribution.
qqnorm(proportion.avian.cameras.long$value)
qqline(proportion.avian.cameras.long$value)

# Check for normality.
shapiro.test(proportion.avian.cameras.long$value)

# Run a t-test.
t.test(proportion.avian.cameras$prop.av.mass, proportion.avian.cameras$prop.av.count)
```

Same result here: means not equal. Finally, proprtion squirrel.

```{r}
# Do the same with proportion avian biomass/counts.

# Calculate biomass and counts for mammalian prey.
proportion.squirrel.cameras <- diet.items %>% 
  filter(method == 'camera') %>% 
  group_by(nest) %>% 
  mutate(total.mass=sum(mass), total.count=n()) %>% 
  filter(genus == 'Tamiasciurus') %>% 
  mutate(mass.sq=sum(mass), count.sq=n(), 
         prop.sq.mass=mass.sq/total.mass, prop.sq.count=count.sq/total.count) %>% 
  select(nest, prop.sq.mass, prop.sq.count) %>% distinct() 

proportion.squirrel.cameras

# Also in long format.
proportion.squirrel.cameras.long <- proportion.squirrel.cameras %>% 
  pivot_longer(!nest, names_to='metric', values_to='value')

# Look at the distribution.
qqnorm(proportion.squirrel.cameras.long$value)
qqline(proportion.squirrel.cameras.long$value)

# Check for normality.
shapiro.test(proportion.squirrel.cameras.long$value)

# Run a t-test.
t.test(proportion.squirrel.cameras$prop.sq.mass, proportion.squirrel.cameras$prop.sq.count)
```

So those p-values aren't enormous, which implies the means are broadly similar. I'm not going to give up using biomass because I think it's more biologically relevant, but I'll have to be sure to include both mass and counts so I can make comparisons with other authors.

The next thing to compare is cameras vs. remains. This cannot be grouped by site because the sample size for the remains + pellets data is too small.

```{r}
# Do the same with proportion avian biomass/counts.

# Calculate biomass for squirrels.
proportion.squirrel.remains <- diet.items %>% 
  filter(method == 'remains') %>% 
  mutate(total.mass=sum(mass, na.rm=TRUE)) %>% 
  filter(genus == 'Tamiasciurus') %>% 
  mutate(mass.sq=sum(mass), prop.sq.mass=mass.sq/total.mass) %>% 
  select(prop.sq.mass) %>% distinct() 

proportion.squirrel.remains

proportion.squirrel.cameras$prop.sq.mass %>% mean()
```

``` {r}
# Calculate biomass for mammals
proportion.mammal.remains <- diet.items %>% 
  filter(method == 'remains') %>% 
  mutate(total.mass=sum(mass, na.rm=TRUE)) %>% 
  filter(class == 'Mammalia') %>% 
  mutate(mass.mm=sum(mass), prop.mm.mass=mass.mm/total.mass) %>% 
  select(prop.mm.mass) %>% distinct() 

proportion.mammal.remains

proportion.mammal.cameras.wide$prop.mm.mass %>% mean()
```

```{r}
# Calculate biomass for birds
proportion.avian.remains <- diet.items %>% 
  filter(method == 'remains') %>% 
  mutate(total.mass=sum(mass, na.rm=TRUE)) %>% 
  filter(class == 'Aves') %>% 
  mutate(mass.av=sum(mass), prop.av.mass=mass.av/total.mass) %>% 
  select(prop.av.mass) %>% distinct() 

proportion.avian.remains

proportion.avian.cameras$prop.av.mass %>% mean()
```

I don't even need to do any statistical tests to see these numbers are radically different. Like, horribly different. Like... I don't even know what to do with these.

And on that depressing note, I think I'll call it quits for the day.







