---
title: "storytime"
output: 
  html_notebook:
    toc: true
---

```{r options, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE)
```

# Introduction

What I did over my summer vacation.

# Goshawk diet

```{r setup}
# Import conflict settings.
source('../src/conflicted.R')

# Load some libraries.
options(tidyverse.quiet = TRUE)
options(dplyr.summarise.inform = FALSE)
library(tidyverse)
library(lubridate)
library(vegan)
library(ggplot2)
library(knitr)
library(kableExtra)

# Bring in data.
source('../src/prey_attributes_revised.R')
```

## Pellets 

```{r}
pellets <- diet.items %>% filter(source == 'P')
```

A total of `r nrow(pellets)` prey items were identified from egested pellets collected from `r distinct(pellets, site) %>% nrow()` sites in 2019 and 2020 (more data from the 2020 season still to come). Most sites had only 1-3 items identified from pellets, though some had 10 or more items identified.

```{r}
pellets %>% group_by(site) %>% summarize(count=n()) %>%
  ggplot(aes(x=count)) +
  geom_histogram(binwidth=1, fill='burlywood3', color='white') +
  labs(x='number of prey items', y='number of sites', title='Prey items per site identified from pellets') +
  theme_classic()
```

`r round(pellets %>% filter(genus != 'Unknown') %>% nrow() / nrow(pellets), digits=2) * 100`% of items could be identified to genus or better. `r pellets %>% filter(binomial != 'Unidentified item') %>% distinct(binomial) %>% nrow()` different species were identified:

`r pellets %>% filter(binomial != 'Unidentified item') %>% distinct(binomial, common) %>% kable() %>% kable_styling()`








