---
title: "Moar specimens"
output: html_notebook
---

```{r options, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, message=FALSE, warning=FALSE)
```

This is the process used to take the specimen data and prep it for further analysis.

```{r message=FALSE, warning=FALSE}
# Import conflict settings.
source('../src/conflicted.R')

# Load some libraries.
library(tidyverse)
library(lubridate)
library(stringr)

# Make a little function for checking at random rows.
grab.some <- function(df, n=10) {
  index <- sample(1:nrow(df), n)
  df[index,]
}

# Bring in the data.
rd <- read_csv('../data/raw/20210118_specimens.csv', guess_max=7000)

# Take a look.
grab.some(rd) %>% select(-c(easting, northing))
```

First off, there are some weird cases I know I need to remove.

* COM is a site from which a few specimens were collected during surveys; however, no NOGO nest or conclusive NOGO sign was detected. This means there's no real reason to assume these are NOGO prey remains and not, say, BADO prey remains.
* Item 1120 was originally identified as SSHA, but later re-IDed as dead NOGO chick. This does not appear to be a genuine prey item  so it should be removed from the analysis.
* Item 1171 was collected as a presumed prey item, but it appears to actually be a shed NOGO feather and so also not actually prey.

```{r}
# Remove problematic items.
rd <- rd %>% 
  filter(id != 1120) %>% 
  filter(id != 1171) %>% 
  filter(site != 'Comsock')
```

Now we start checking for typos.

```{r}
# Check for typos in site names.

rd %>% distinct(site)
```

Great, but let's rename that, bring in abbreviations, and make sure everything matches.

```{r message=FALSE}
# Bring in a list of site abbreviations and site names.
nest.list <- read_csv('../data/processed/site_abbreviations.csv')

rd %>% rename(name=site) %>% left_join(nest.list, by=c('name')) %>% 
  distinct(site, name)
```

Uh-huh so Gowan and Glacial Lake didn't join properly.

```{r}
# Rename problems.
rd$site <- str_replace_all(rd$site, c('Gowan'='Gowan Creek', 'Glacial Lake'='Glacier Lake'))

# Join again.
rd <- rd %>% rename(name=site) %>% left_join(nest.list, by=c('name'))
```

Check again.

```{r}
distinct(rd, site, name)
```

Check for typos in class.

```{r}
distinct(rd, class)
```

Great. Order?

```{r}
distinct(rd, order)
```

Great. Family?

```{r}
distinct(rd, family)
```

Great. Genus?

```{r}
distinct(rd, genus)
```

Not precisely a typo, but I used the wrong genus (Picoides) for hairy and downy woodpeckers. There appears to be some confusion on whether it's Picoides, Leuconotopicus, or Dryobates, but BOW
uses Dryobates.

```{r}
# Fix out-of-date taxonomy.
rd$genus <- str_replace(rd$genus, 'Picoides', 'Dryobates')

# Check again.
distinct(rd, genus)
```

Great. On to species.

```{r}
distinct(rd, species)
```

Great. Common names?

```{r}
distinct(rd, common)
```

Ok, so some problems here in that I shouldn't have "grouse" and "duck," but I'll fix that later in the more complex cleanup stage.

Finally, age and size.

```{r}
distinct(rd, age)

distinct(rd, size)
```

Perfect.  And I would like to replace all the Us with "Unknown" for taxonomy.

```{r message=FALSE, warning=FALSE}
# Replace Us.
rd <- rd %>% mutate_at(c('class', 'order', 'family', 'genus', 'species'), funs(case_when(
    . == 'U' ~ 'Unknown',
    TRUE ~ .
  )))

# Then make species lowercase.
rd$species <- str_to_lower(rd$species)
```

Now let's look for oddities.

```{r}
rd %>% group_by(class, order, family, genus, species, common, size, age) %>% 
  mutate(count=n()) %>% 
  distinct(class, order, family, genus, species, common, size, age, count) %>% 
  arrange(class, order, family, genus, species, common, size)
```

Numerous problems:

* I don't like the ducks as both Anas sp. and Anatidae. Probably the unidentifed Anatidae is an Anas (it's that tricky feather I couldn't find the right blue-green match to), so I can make that Anas sp.
* The unknown grouse can be given their own biomass.
* There's really ony one Catharus option, so I should drop the sp.
* All the voles should be given vole mass, not small mammal mass.
* I said the Dryobates was M... and downy woodpeckers are small... so it must, therefore, be a hairy.


```{r}
rd <- rd %>% mutate(species=case_when(
  genus == 'Anas' & species == 'sp' ~ 'unknown',
  genus == 'Catharus' ~ 'ustulatus',
  genus == 'Dryobates' & species == 'sp' ~ 'villosus',
  TRUE ~ species
))

rd <- rd %>% mutate(common=case_when(
  genus == 'Catharus' ~ 'swainson\'s thrush',
  family == 'Phasianidae' & species == 'unknown' ~ 'grouse',
  genus == 'Anas' & species == 'unknown' ~ 'duck',
  genus == 'Dryobates' & species == 'villosus' ~ 'hairy woodpecker',
  family == 'Anatidae' & genus == 'Unknown' ~ 'unknown', 
  TRUE ~ common
))
```

Check one more time.

```{r}
rd %>% group_by(class, order, family, genus, species, common, size, age) %>% 
  mutate(count=n()) %>% 
  distinct(class, order, family, genus, species, common, size, age, count) %>% 
  arrange(class, order, family, genus, species, common, size)
```

And let's see how well the mass numbers join.

```{r message=FALSE}
# Bring in a list of all known prey.
prey.list <- read_csv('../data/interim/prey_attributes.csv')

# Join the biomass data to the list of diet items.
ri <- prey.list %>% select(genus, species, binomial, common, category, mass) %>% 
  right_join(rd, by=c('genus', 'species', 'common'))

# And check.
ri %>% distinct(class, order, family, genus, species, common, mass) %>% 
  arrange(class, order, family, genus, species, common)
```

Looks good. Let's save it to cut down on processing later.

```{r}
write_csv(rd, '../data/processed/specimens_20210318.csv')
```








