---
title: "Clean Cameras"
output: html_notebook
---

```{r options, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, message=FALSE, warning=FALSE)
```

This is the process used to take the raw camera data produced by my amazing minion and prep it for further analysis.

```{r message=FALSE, warning=FALSE}
# Import conflict settings.
source('../src/conflicted.R')

# Load some libraries.
library(tidyverse)
library(lubridate)
```

All files, and *only* files, containing camera data contain the word "photo," which makes things much, much easier.

```{r message=FALSE}
# Get a list of all the camera data files.
file.list <- list.files(path='../data/raw', pattern='photos', full.names=TRUE)

# Open all the files in the list and add to a data frame.
df <- file.list %>% map(read_csv, col_types = cols(.default = "c")) %>% 
  bind_rows()
```

There are some horrible, horrible problems with time and date formats, and sadly I think the only real way to fix them is to go through the sheets one by one and standardize them. That is definitely a thing I will have to do at some point, but thankfully for this analysis I only need year and the date of the first photo.

```{r}
# Try to fix all of the different datetime formats.
df$datetime <- parse_date_time(df$datetime, c('ymd HMS', 'mdy HMS', 'dmy HMS', 'mdy HM'))
```

Now on to hunting down and fixing typos, etc.

```{r}
# Typos in the "interest" column.
df %>% distinct(interest)
```

Yes, some.

```{r}
# Fix the two typos in 'interest'.
df <- df %>% mutate(interest=case_when(
  interest == 'Delivery' ~ 'delivery',
  interest == 'preu' ~ 'prey',
  TRUE ~ interest
))

# Typos in 'class'?
df %>% distinct(class)
```

No, that looks fine, but I would like to replace the Us with Unknowns.

```{r}
# Replace Us with Unknown.
df <- df %>% mutate(class=case_when(
  class == 'U' ~ 'Unknown',
  TRUE ~ class
)) 

# Typos in 'order'?
df %>% distinct(order)
```

A few. Fix them, and again turn Us to Unknowns

```{r}
df <- df %>% mutate(order=case_when(
  order == 'Passerofprmes' ~ 'Passeriformes',
  order == 'Passiformes' ~ 'Passeriformes',
  order == 'U' ~ 'Unknown',
  TRUE ~ order
))
```

On to family.


```{r}
df %>% distinct(family)
```

Yes, a number of them genera mistakenly entered into the family column.

```{r}
# Fix the typos in 'family.'
df <- df %>% mutate(family=case_when(
  family == 'Sciruidae' ~ 'Sciuridae',
  family == 'Lepus' ~ 'Leporidae',
  family == 'Dendragapus' ~ 'Phasianidae',
  family == 'Neotoma' ~ 'Cricetidae',
  family == 'U' ~ 'Unknown',
  TRUE ~ family
))

# Typos in 'genus'?
df %>% distinct(genus)
```

Several. I'm using *Dryobates* instead of *Leuconotopicus* for the genus of HAWO and DOWO. 

There's two tricky ones here, *Tamias* which might be *Neotamias* or *Tamiasciurus*, and *Phasianidae*, which might be *Bonasa* or *Dendragapus*.

```{r}
# What's the squirrel?
df %>% filter(genus == 'Tamias') %>% select(genus, species)

# What's the squirrel?
df %>% filter(genus == 'Tamias') %>% select(genus, species)

# What's the grouse?
df %>% filter(genus == 'Phasianidae') %>% select(genus, species)
```

```{r}
# Fix the typos.
df <- df %>% mutate(genus=case_when(
  genus == 'Tamasciurus' ~ 'Tamiasciurus',
  genus == 'Tamasciuris' ~ 'Tamiasciurus',
  genus == 'Leuconotopicus' ~ 'Dryobates',
  genus == 'Turdidae' ~ 'Turdus',
  genus == 'Leporidae' ~ 'Lepus',
  genus == 'Tamias' ~ 'Neotamias',
  genus == 'Phasianidae' ~ 'Dendragapus',
  genus == 'Cricetidae' ~ 'Neotoma',
  genus == 'U' ~ 'Unknown',
  TRUE ~ genus
))

# Typos in species?
df %>% distinct(species)
```

Remarkably free of typos, just a few things to fix up.

```{r}
# Fix the typos in 'species'.
df <- df %>% mutate(species=case_when(
  species == 'spp' ~ 'sp',
  species == 'U' ~ 'unknown',
  TRUE ~ species
))

# How about common names?
df %>% distinct(common)
```

Kind of a hot mess. Exactly what you'd expect from common names, really. Let's do this in stages.

The tricky one is the one that's just 'm'.

```{r}
# Make everything lowercase.
df$common <- df$common <- str_to_lower(df$common)

# Split on the periods to make separate words.
df$common <- gsub('\\.', ' ', df$common)

# Find the problem children.
df %>% filter(common == 'm') %>% select(genus, species, common)

# Fix the typos.
df <- df %>% mutate(common=case_when(
  common == 'bad-tailed pigeon' ~ 'band-tailed pigeon',    # HAHAHAHA I love this one!
  common == 'chimpunk' ~ 'chipmunk',
  common == 'stellers jay' ~ 'steller\'s jay',
  common == 'townsends chipmunk' ~ 'townsend\'s chipmunk',
  common == 'm' & species == 'douglasii' ~ 'douglas squirrel',
  common == 'm' & genus == 'Rattus' ~ 'rat',
  common == 'u' ~ 'unknown',
  TRUE ~ common
))

# Lastly, see if anything is wonky in 'size'.
df %>% distinct(size)
```

Looks perfect.

In the earliest data sheets I forgot to include a column for 'order' because I'm an idiot. Not that things are cleaned up I can fill in those missing data.

```{r message=FALSE}
# Bring in a list of all known prey.
prey.list <- read_csv('../data/interim/prey_attributes.csv')

# Join the camera data to the list of diet items.
df <- prey.list %>% select(order.list=order, genus, species, common) %>% 
  right_join(df, by=c('genus', 'species', 'common')) %>% 
  arrange(site, datetime) %>%  
  mutate(order=case_when(
    is.na(order) ~ order.list,
    TRUE ~ order
  ))
```

A final check.

```{r}
df %>% select(class, order, family, genus, species, common) %>% distinct() %>% 
  arrange(class, order, family, genus, species, common)
```

This turns up one last typo: a known genus (Turdus) with an unknown family. Otherwise, I can fill all the NAs with 'Unknown'--except, of course, for those rows that are nothing at all.

```{r warning=FALSE}
# Fill in the last typo.
df <- df %>% mutate(family=case_when(
  genus == 'Turdus' & is.na(family) ~ 'Turdidae',
  TRUE ~ family
))

# Fill in the NAs.
df <- df %>% mutate_at(c('order', 'family', 'genus'), funs(case_when(
  is.na(.) & class != 'Unknown' ~ 'Unknown',
  TRUE ~ .
)))
```

And that leads me to the final, but question: what to do with half-identified things??

Problems include unidentifed:

* grouse
* thrushes
* tree squirrels
* chipmunks

There's only one townsend's chipmunk, so I can easily assign that to *Neotamias sp.*. Likewise there's a small number *Tamiasciurus sp.* so I can assign them to *T. douglasii*.



