---
title: "Clean Cameras"
output: html_notebook
---

```{r options, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, message=FALSE, warning=FALSE)
```

This is the process used to take the raw camera data produced by my amazing minion and prep it for further analysis.

```{r message=FALSE, warning=FALSE}
# Import conflict settings.
source('../src/conflicted.R')

# Load some libraries.
library(tidyverse)
library(lubridate)
```

All files, and *only* files, containing camera data contain the word "photo," which makes things much, much easier.

```{r message=FALSE}
# Get a list of all the camera data files.
file.list <- list.files(path='../data/raw', pattern='photos', full.names=TRUE)

# Open all the files in the list and add to a data frame.
df <- file.list %>% map(read_csv, col_types = cols(.default = "c")) %>% 
  bind_rows()
```

There are some horrible, horrible problems with time and date formats, and sadly I think the only real way to fix them is to go through the sheets one by one and standardize them. That is definitely a thing I will have to do at some point, but thankfully for this analysis I only need year and the date of the first photo.

```{r}
# Try to fix all of the different datetime formats.
df$datetime <- parse_date_time(df$datetime, c('ymd HMS', 'mdy HMS', 'dmy HMS', 'mdy HM'))
```

Now on to hunting down and fixing typos, etc.

```{r}
# Typos in the "interest" column.
df %>% distinct(interest)
```

Yes, some.

```{r}
# Fix the two typos in 'interest'.
df <- df %>% mutate(interest=case_when(
  interest == 'Delivery' ~ 'delivery',
  interest == 'preu' ~ 'prey',
  TRUE ~ interest
))

# Typos in 'class'?
df %>% distinct(class)
```

No, that looks fine, but I would like to replace the Us with Unknowns.

```{r}
# Replace Us with Unknown.
df <- df %>% mutate(class=case_when(
  class == 'U' ~ 'Unknown',
  TRUE ~ class
)) 

# Typos in 'order'?
df %>% distinct(order)
```

A few. Fix them, and again turn Us to Unknowns

```{r}
df <- df %>% mutate(order=case_when(
  order == 'Passerofprmes' ~ 'Passeriformes',
  order == 'Passiformes' ~ 'Passeriformes',
  order == 'U' ~ 'Unknown',
  TRUE ~ order
))
```

On to family.


```{r}
df %>% distinct(family)
```

Yes, a number of them genera mistakenly entered into the family column.

```{r}
# Fix the typos in 'family.'
df <- df %>% mutate(family=case_when(
  family == 'Sciruidae' ~ 'Sciuridae',
  family == 'Lepus' ~ 'Leporidae',
  family == 'Dendragapus' ~ 'Phasianidae',
  family == 'Neotoma' ~ 'Cricetidae',
  family == 'U' ~ 'Unknown',
  TRUE ~ family
))

# Typos in 'genus'?
df %>% distinct(genus)
```

Several. I'm using *Dryobates* instead of *Leuconotopicus* for the genus of HAWO and DOWO. 

There's two tricky ones here, *Tamias* which might be *Neotamias* or *Tamiasciurus*, and *Phasianidae*, which might be *Bonasa* or *Dendragapus*.

There's also some new genera since I last ran this. One looks to be mispelled, a couple I'm just curious about.

```{r}
# What's the squirrel?
df %>% filter(genus == 'Tamias') %>% select(genus, species)

# What's the grouse?
df %>% filter(genus == 'Phasianidae') %>% select(genus, species)

# What's the weird typo-looking one?
df %>% filter(genus == 'Oxpreus') %>% select(class, family, genus, species)
# Oh, a varied thrush.

# And what are the new ones?
df %>% filter(genus %in% c('Bubo', 'Mustela', 'Marmota'))

# What's up with the accipiters?
df %>% filter(order == 'Accipitriformes')
```
Hot damn! Great-horned owl and a weasel and a marmot!! How cool!

Ok, looking closer at the marmot I have to say I'm unconvinced. It's definitely a big mammal, but it just looks so black-and-white and stripey, and a marmot is just so... huge. It's about 3 times the weight of a hare, and a hare is already a pretty big mouthful for a NOGO! So I think I'll just make this an unknown large mammal.

```{r}
# Fix the typos.
df <- df %>% mutate(genus=case_when(
  genus == 'Tamasciurus' ~ 'Tamiasciurus',
  genus == 'Tamasciuris' ~ 'Tamiasciurus',
  genus == 'Leuconotopicus' ~ 'Dryobates',
  genus == 'Turdidae' ~ 'Turdus',
  genus == 'Patagoienas' ~ 'Patagioenas',
  genus == 'Leporidae' ~ 'Lepus',
  genus == 'Tamias' ~ 'Neotamias',
  genus == 'Phasianidae' ~ 'Dendragapus',
  genus == 'Cricetidae' ~ 'Neotoma',
  genus == 'Drycopus' ~ 'Dryocopus',
  genus == 'U' ~ 'Unknown',
  genus == 'Oxpreus' ~ 'Ixoreus',
  TRUE ~ genus
))

# Change the marmot.
df <- df %>% mutate(order=case_when(
  species == 'caligata' ~ 'Unknown',
  TRUE ~ order
),
                    family=case_when(
  species == 'caligata' ~ 'Unknown',
  TRUE ~ family
                    ),
                    genus=case_when(
  species == 'caligata' ~ 'Unknown',
  TRUE ~ genus
                    ),
                    common=case_when(
  species == 'caligata' ~ 'unknown',
  TRUE ~ common
                    ),
                    species=case_when(
  species == 'caligata' ~ 'unknown',
  TRUE ~ species
                    ))

# Typos in species?
df %>% distinct(species)
```

Remarkably free of typos, just a few things to fix up.

```{r}
# Fix the typos in 'species'.
df <- df %>% mutate(species=case_when(
  species == 'fulignosus' ~ 'fuliginosus',
  genus == 'Sorex' ~ 'sp',
  genus == 'Myotis' ~ 'sp',
  species == 'spp' ~ 'sp',
  species == 'U' ~ 'unknown',
  TRUE ~ species
))

# How about common names?
df %>% distinct(common)
```

Kind of a hot mess. Exactly what you'd expect from common names, really. Let's do this in stages.

The tricky one is the one that's just 'm'.

```{r}
# Make everything lowercase.
df$common <- df$common <- str_to_lower(df$common)

# Split on the periods to make separate words.
df$common <- gsub('\\.', ' ', df$common)

# Find the problem children.
df %>% filter(common == 'm') %>% select(genus, species, common)

# Fix the typos.
df <- df %>% mutate(common=case_when(
  common == 'bad-tailed pigeon' ~ 'band-tailed pigeon',    # HAHAHAHA I love this one!
  common == 'chimpunk' ~ 'chipmunk',
  common == 'northern flying squirrel' ~ 'flying squirrel',
  common == 'stellers jay' ~ 'steller\'s jay',
  common == 'swainsons thrush' ~ 'swainson\'s thrush',
  common == 'townsends chipmunk' ~ 'townsend\'s chipmunk',
  common == 'm' & species == 'douglasii' ~ 'douglas squirrel',
  common == 'm' & genus == 'Rattus' ~ 'rat',
  genus == 'Sorex' ~ 'shrew',
  genus == 'Myotis' ~ 'bat',
  common == 'stoat' ~ 'short-tailed weasel',
  common == 'u' ~ 'unknown',
  TRUE ~ common
))

# Lastly, see if anything is wonky in 'size'.
df %>% distinct(size)
```

Looks perfect.

In the earliest data sheets I forgot to include a column for 'order' because I'm an idiot. Now that things are cleaned up I can fill in those missing data.

```{r message=FALSE}
# Bring in a list of all known prey.
prey.list <- read_csv('../data/interim/prey_attributes.csv')

# Join the camera data to the list of diet items.
df <- prey.list %>% select(order.list=order, genus, species, common) %>% 
  right_join(df, by=c('genus', 'species', 'common')) %>% 
  arrange(site, datetime) %>%  
  mutate(order=case_when(
    is.na(order) ~ order.list,
    TRUE ~ order
  ))
```

A final check.

```{r}
df %>% select(class, order, family, genus, species, common) %>% distinct() %>% 
  arrange(class, order, family, genus, species, common)
```

This turns up a few last typos: a known genus (Turdus) with an unknown family, and hare is marked as a rodent. Also there's an unknown Accipitridae that could be a red-tailed hawk, so I'll just call it a RTHA. Otherwise, I can fill all the NAs with 'Unknown'--except, of course, for those rows that are nothing at all.

```{r warning=FALSE}
# Fill in the last typos.
df <- df %>% mutate(family=case_when(
  genus == 'Turdus' & is.na(family) ~ 'Turdidae',
  TRUE ~ family
),
order=case_when(
  genus == 'Lepus' ~ 'Lagomorpha',
  TRUE ~ order
))

# Change the unknown accipiter to RTHA.
df <- df %>% 
  mutate(species=case_when(family == 'Accipitridae' & genus == 'Unknown' ~ 'jamaicensis',
                               TRUE ~ species),
              common=case_when(family == 'Accipitridae' & genus == 'Unknown' ~ 'red-tailed hawk',
                              TRUE ~ common),
              genus=case_when(family == 'Accipitridae' & genus == 'Unknown' ~ 'Buteo',
                                TRUE ~ genus))

# Fill in the NAs.
df <- df %>% mutate_at(c('order', 'family', 'genus'), funs(case_when(
  is.na(.) & class != 'Unknown' ~ 'Unknown',
  TRUE ~ .
)))
```

And that leads me to the final, but question: what to do with half-identified things??

Problems include unidentified:

* grouse
* chipmunks
* tree squirrels
* thrushes

The chipmunk is actually the inverse problem, only one has been IDed down to species. I can easily make the *Neotamias townsendii* a *N. sp.* Likewise I can change all the *Tamiasciurus sp* to *Tamiasciurus douglasii*. The unknown *Turdus* is actually another typo, because *T. americanus* is the only species present. 

So that just leaves the grouse. When looking at the combined camera and physical data, *Dendragapus* appears more common. So I guess I can assign my unknown grouse to that, since there's only one of them.

```{r}
df <- df %>% mutate(species=case_when(
  genus == 'Tamiasciurus' & species == 'unknown' ~ 'douglasii',
  genus == 'Neotamias' & species == 'townsendii' ~ 'sp',
  genus == 'Turdus' & species == 'unknown' ~ 'migratorius',
  family == 'Phasianidae' & species == 'unknown' ~ 'fuliginosus',
  TRUE ~ species
              ),
              genus=case_when(
  family == 'Phasianidae' & species == 'fuliginosus' ~ 'Dendragapus',
  TRUE ~ genus
              ),
              common=case_when(
  genus == 'Tamiasciurus' & species == 'douglasii' ~ 'douglas squirrel',
  genus == 'Neotamias' ~ 'chipmunk',
  genus == 'Turdus' ~ 'american robin',
  genus == 'Dendragapus' ~ 'sooty grouse',
  TRUE ~ common
              ))
```

Finally, make sure all of the species and sizes match up.

```{r}
df %>% select(class, order, family, genus, species, common, size) %>% distinct() %>% 
  arrange(class, order, family, genus, species, common, size)
```

Ah, yes, a few errors here. Are any of them actually correct?

```{r}
# Are the small grouse chicks?
df %>% filter(family == 'Phasianidae' & size == 'S')
# Yes. Leave them alone.

# And the steller's jays?
df %>% filter(genus == 'Cyanocitta' & size == 'S')
# Yes, also juveniles.

# Robins?
df %>% filter(genus == 'Turdus' & size == 'S')
# Yes, also juveniles.

# Woodrats?
df %>% filter(genus == 'Neotoma' & size == 'S')
# No, that's a typo, should be M.

# Rats?
df %>% filter(genus == 'Rattus' & species == 'sp')
```

That turns up an interesting one. *Rattus sp.* should be medium, since all the possible rat species are about that size. But there's something here marked as *R. sp* that might be a mountain beaver?? I think I'll just call that one a mountain beaver!

```{r}
# Change from unknown rat to mountain beaver.
df <- df %>% 
  mutate(family=case_when(genus == 'Rattus' & !is.na(comments) & size == 'L' ~ 'Aplodontiidae',
                               TRUE ~ family),
              genus=case_when(genus == 'Rattus' & !is.na(comments) & size == 'L' ~ 'Aplodontia',
                              TRUE ~ genus),
              species=case_when(genus == 'Aplodontia' & !is.na(comments) & size == 'L' ~ 'rufa',
                                TRUE ~ species),
              common=case_when(genus == 'Aplodontia' & !is.na(comments) & size == 'L' ~ 'mountain beaver',
                               TRUE ~ common))

# Let's keep going.

# What's up with the unsized squirrel?
df %>% filter(genus == 'Tamiasciurus' & is.na(size))
# Just a typo.

# And the new great-horned owl.
df %>% filter(genus == 'Bubo')
# Yes, it's a juvenile.

# Ok, let's fix all of these.
df <- df %>% mutate(size=case_when(
  genus == 'Catharus' ~ 'S',
  species == 'pubescens' ~ 'S',
  species == 'villosus' ~ 'M',
  genus == 'Lepus' & is.na(size) ~ 'L',
  genus == 'Neotoma' ~ 'M',
  genus == 'Rattus' ~ 'M',
  genus == 'Neotamias' ~ 'S',
  genus == 'Tamiasciurus' & is.na(size) ~ 'M',
  TRUE ~ size
))
```

I haven't been using age in my analyses, but on further consideration I probably should... there's more juveniles here than I realized. That means I need some way to mark which items are juveniles or not.

```{r}
# Mark things with 'juvenile' in the comments as J, otherwise U for unknown.
df <- df %>% mutate(age=case_when(
  str_detect(comments, 'juv|Juv') ~ 'J',
  str_detect(comments, 'baby|Baby') & !is.na(class) ~ 'J',
  TRUE ~ 'U'
  
))
```

Okay, and...

```{r}
# And one last check.
df %>% select(class, order, family, genus, species, common, size) %>% distinct() %>% 
  arrange(class, order, family, genus, species, common, size)
```

I think that's... done! Now let's make it pretty and save it.

```{r}
# Tidy up the data set.
df <- df %>% select(filename, serial, datetime, 
              site, interest, live.chicks, 
              class, order, family, genus, species, common,
              size, age, comments, sex) %>% 
  arrange(site, datetime)

# Save as a csv.
write_csv(df, '../data/interim/cameras_20210623.csv')
```

