---
title: "Finalish cameras"
output: github_document
---

```{r options, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, message=FALSE, warning=FALSE)
```

I think I have all the data I need to run the productivity ~ diet models. So here goes...

```{r message=FALSE, warning=FALSE}
# Import conflict settings.
source('../src/conflicted.R')

# Load some libraries.
library(tidyverse)
library(lubridate)
library(ggplot2)
library(vegan)
library(broom.mixed)
library(lme4)

# Bring in camera data.
camera.data <- read_csv('../data/interim/cameras_20210315.csv', guess_max=7000) %>% 
  ## filter only records with at least size assigned...
  filter(size != 'U')

# To make sure the average sizes work, I also need to bring in the physical remains.
remains.data <- read_csv('../data/raw/20210118_specimens.csv', guess_max=7000) %>% 
  ## filter only records with at least size assigned...
  filter(size != 'U') %>% 
  ## and rename a confusing column.
  rename(name=site) %>% 
  ## and fix a typo
  mutate(genus=case_when(
    genus == 'Picoides' ~ 'Dryobates',
    TRUE ~ genus
  ))

# Bring in a list of site abbreviations and site names.
nest.list <- read_csv('../data/processed/site_abbreviations.csv')

# Standardize the specimen data.
remains.data <- left_join(remains.data, nest.list, by='name') %>% 
  filter(id != 1120) %>% 
  filter(id != 1171) %>% 
  filter(site != 'COM') %>% 
  select(site, date, class, order, family, genus, species, common, size, age, source)

# Standardize the camera data.
camera.data <- camera.data %>% 
  mutate(date=date(datetime), source='C') %>% 
  select(site, date, class, order, family, genus, species, common, size, age, source)

# Join them together.
diet.data <- bind_rows(remains.data, camera.data)

# Add a unique site/year identifier.
diet.data <- diet.data %>% 
  mutate(year=year(date), nest=paste(site, year, sep=''))

# Do some cleanup.
diet.data <- diet.data %>% 
  mutate(size=case_when(
    size %in% c('S', 'Small') ~ 'small',
    size %in% c('M', 'Medium') ~ 'medium',
    size %in% c('L', 'Large') ~ 'large'
  )) %>%  
  ## and replace all the "U"s with "Unknown"...
  mutate_at(c('class', 'order', 'family', 'genus'), funs(case_when(
    . == 'U' ~ 'Unknown',
    TRUE ~ .
  ))) %>% 
  ## in species as well...
  mutate(species=case_when(
    species == 'U' ~ 'unknown',
    TRUE ~ species
  ))
```

That's the diet data from cameras brought in. Time to add the mass data.

```{r message=FALSE}
# Bring in a list of all known prey.
prey.list <- read_csv('../data/interim/prey_attributes.csv')

# Join the biomass data to the list of diet items.
diet.items <- prey.list %>% select(genus, species, binomial, common, category, mass) %>% 
  right_join(diet.data, by=c('genus', 'species', 'common'))

# For unidentified items, classify them by size and class.
diet.items <- diet.items %>% mutate(category=case_when(
  is.na(category) & class == 'Mammalia' & size == 'small' ~ 'small mammal',
  is.na(category) & class == 'Mammalia' & size == 'medium' ~ 'medium mammal',
  is.na(category) & class == 'Mammalia' & size == 'large' ~ 'large mammal',
  is.na(category) & class == 'Aves' & size == 'small' ~ 'small bird',
  is.na(category) & class == 'Aves' & size == 'medium' ~ 'medium bird',
  is.na(category) & class == 'Aves' & size == 'large' ~ 'large bird',
  is.na(category) & class == 'Unknown' ~ paste(size, 'item'),
  TRUE ~ category))

# For unidentified items, fill in the binomial column.
diet.items <- diet.items %>% replace_na(list(binomial = 'Unidentified item'))

# Calculate average masses for unidentified items, based of known species.
mean.mass <- diet.items %>% drop_na(mass) %>% 
  distinct(binomial, mass, category) %>% 
  group_by(category) %>% 
  ## averaging the mass for each size & class category...
  summarize(average=mean(mass)) %>% 
  pivot_wider(names_from=category, values_from=average) %>% 
  ## calculating the average mass for complete unknowns...
  mutate(`large item` = mean(`large bird`, `large mammal`),
         `medium item` = mean(`medium bird`, `medium mammal`),
         `small item` = mean(`small bird`, `small mammal`)) %>% 
  ## and reassembling it in a tidy format.
  pivot_longer(everything(), names_to='category', values_to='average')

# Join average mass to diet items...
diet.items <- left_join(diet.items, mean.mass, by='category') %>% 
  ## and fill in missing mass with average values
  mutate(mass=coalesce(mass, average)) %>% 
  ## then drop no longer needed average column and rearrange.
  select(site, year, nest, class, order, family, genus, species, binomial, common, 
         category, size, age, mass, source)
```

Looks good. One last thing to change is that juveniles need to have smaller mass than adults. This is a completely arbitrary choice. I don't think I have any small species juveniles, only large and mediums. Miller uses 65-95% of adults mass for medium birds and 55-95% for large birds, and 50% for mammals. That seems incredibly fine considering the coarseness of my data, so maybe I'll just go with a flat 50%?

```{r}
# Change mass for juvenile items.
diet.items <- diet.items %>% mutate(mass=case_when(
  age == 'J' ~ 0.5*mass,
  TRUE ~ mass
))
```

Ok, now on to doing all the calculations needed to get things ready for the models.

```{r}
# Make a function to calculate class mass.
class.mass <- function(data) {
    data %>% mutate(total.mass=sum(.data$mass)) %>%
    filter(class %in% c('Mammalia', 'Aves')) %>%
    group_by(class) %>% 
    mutate(class.mass=sum(.data$mass), per.class=class.mass/total.mass*100) %>% 
    distinct(class, per.class) %>% 
    rename(variable=class, percent=per.class)
}

# And a function to calculate squirrel mass.
squirrel.mass <- function(data) {
    data %>% mutate(total.mass=sum(.data$mass)) %>%
    filter(genus == 'Tamiasciurus') %>%
    mutate(genus.mass=sum(.data$mass), per.genus=genus.mass/total.mass*100) %>% 
    distinct(genus, per.genus) %>% 
    rename(variable=genus, percent=per.genus)
}

# Calculate percent mass for each nest.
nest.mass <- diet.items %>% filter(source == 'C') %>% 
  group_by(nest) %>% nest() %>% 
  mutate(class=map(data, class.mass), 
         genus=map(data, squirrel.mass)) %>% 
  pivot_longer(-c(nest, data), names_to='var', values_to='per') %>% 
  unnest(per) %>% 
  select(nest, variable, percent) %>% 
  pivot_wider(id_cols=c(nest), names_from=variable, values_from=percent, values_fill=0)

# Calculate counts of items identified to genus for each nest.
nest.count <- diet.items %>% filter(source == 'C' & binomial != 'Unidentified item') %>% 
  group_by(nest, binomial) %>% 
  mutate(count=n()) %>% 
  select(nest, binomial, count) %>% 
  distinct() %>% 
  pivot_wider(names_from=binomial, values_from=count, values_fill=list(count = 0))

# Diversity per nest.
nest.diversity <- plyr::ddply(nest.count, ~nest, function(x) {
  data.frame(diet.diversity=diversity(x[-1], index='simpson'))
})
```

Calculations done, wrangle it all into a tidy data set.

```{r message=FALSE}
# Bring in productivity data.
productivity <- read_csv('../data/raw/productivity.csv')

# Make a happy data set for the models to use.
model.data <- productivity %>% mutate(nest=paste(site, year, sep='')) %>% 
  left_join(nest.diversity, by=c('nest')) %>% 
  left_join(nest.mass, by=c('nest')) %>%
  select(site, year, nest, n.fledge, diet.diversity, Tamiasciurus) %>% 
  filter_all(all_vars(!is.na(.)))

model.data
```

PNC looks wrong because the diversity is 0, but that is technically correct: the only things identified with any level of precision were squirrels.

Before we make the models, let's look at some graphs.

```{r}
# All the data together.
ggplot(model.data, aes(x=Tamiasciurus, y=n.fledge, color=site, shape=as.factor(year))) +
  geom_point(size=4) +
  theme_classic()
```

And break it apart to make it a little easier to see.

```{r}
# 2019.
model.data %>% filter(year == 2019) %>% 
  ggplot(aes(x=Tamiasciurus, y=n.fledge, color=site)) +
  geom_point(size=4) +
  ylim(0, 3) +
  theme_classic()

# 2020.
model.data %>% filter(year == 2020) %>% 
  ggplot(aes(x=Tamiasciurus, y=n.fledge, color=site)) +
  geom_point(size=4) +
  ylim(0, 3) +
  theme_classic()
```

Let's try making some models.

```{r}
# Start with just squirrels.
squirrel.model <- lmer(n.fledge ~ Tamiasciurus + (1|year), data=model.data)

# Look at it.
summary(squirrel.model)

# Look at some diagnostics.
data.frame(predicted=predict(squirrel.model, type='response'),
           residuals=residuals(squirrel.model, type='response')) %>% 
  ggplot(aes(x=predicted, y=residuals)) +
  geom_point() +
  geom_hline(yintercept=0, linetype='dashed') +
  geom_smooth(method='lm', formula='y~x', se=FALSE, color='black') +
  theme_classic()
```

That's how well the model is *doing*. But how well is the model *modeling?*